---
title: "Maps with buffers"
author: "Orvahlo Augusto, Joe Brew, Carlos Chaccour, Edgar Jamisse, Charfudin Sacoor, Francisco Saute"
fig_width: 4
output:
  html_document:
    theme: flatly
    toc: yes
    toc_float: yes
  pdf_document:
    toc: yes
fig_height: 2.6
---
```{r, echo = FALSE, warning = FALSE, message = FALSE, comment = NA, error= FALSE, cache = FALSE}
# No scientific notation
options(scipen=999)

# Packages 
library(xtable)
library(knitr)
library(ggplot2) 
# library(dplyr)
library(rgeos)
library(RColorBrewer)
library(Rmisc)
# Libraries
library(rgdal)
library(raster)
library(sp)
library(leaflet)
library(ggmap)
library(readxl)
library(dplyr)
library(readr)
library(geosphere)

# Basic knitr options
opts_chunk$set(comment = NA, 
               echo = FALSE, 
               warning = FALSE, 
               message = FALSE, 
               error = TRUE, # Render report, even with errors
               cache = F)

# Helpers
source('lib/get_zambezia.R')
source('lib/helpers.R')
```

```{r setup, include=FALSE, echo = FALSE}
library(knitr)
knit_engines$set(asis = function(options) {
  if (options$echo && options$eval) knit_child(text = options$code)
})
```


```{r}
#### READ IN AND CLEAN UP DATA

# Read in final dataset, as given by Orvahlo on October 24, 2016
# census <- read_excel('Limpesa_Final__241016_For maps.xlsx')
# Read in corrected final dataset, sent by Edgar on October 30, 2016
# census <- read_excel('Casos problematticos resolvidos_Edgar Jamisse_Mopeia_2016_28_10.xls')
census <- read_csv('Casos problematticos resolvidos_Edgar Jamisse_Mopeia_2016_28_10.csv')
# Fix weird header
names(census)[6] <- 'locality_Final'

# Get more nicely named geographic coordinate column names
census$lat <- census$y <- census$latitude <- census$gpc_lat
census$lon <- census$lng <- census$longitude <-census$x <- census$gpc_lng

# Get a shapefile for Mozambique
moz <- raster::getData('GADM', country = 'MOZ', level = 3)
moz2 <- raster::getData('GADM', country = 'MOZ', level = 2)

# GET CORRECT FINAL COLUMNS, AS EXPLAINED IN EMAIL
# Get a village number
census$village_number <- census$`village number_final`
census$administrative_post <- census$administrative_post_final
census$house_number <- census$houseno_Final_1
census$locality <- census$locality_Final

# Convert village number to numeric
census$village_number <- 
  as.numeric(as.character(census$village_number))

# Remove all NA rows
census <- census %>%
  filter(!is.na(lng),
         !is.na(lat),
         !is.na(village_number),
         !is.na(house_number))


# Save old version of census
census_old <- census

# Create a spatial version of census
census_spatial <- census
coordinates(census_spatial) <- ~x+y
proj4string(census_spatial) <- proj4string(moz)

# Subset the shapefile just to Zambezia
zam <- moz[moz@data$NAME_1 == 'Zambezia',]

# Subset further to just Mopeia
mop <- moz2[moz2@data$NAME_2 == 'Mopeia',]

# Save the original proj4string
original_proj4string <- proj4string(moz)

# Prepare ggmap-compatible objects
zam_fortified <- fortify(zam, id = ID_3)
mop_fortified <- fortify(mop, id = ID_3)
zam_gg <- get_map(location = c(lon = mean(census$lon, na.rm = TRUE),
                               lat = mean(census$lat, na.rm = TRUE)),
                  maptype = 'satellite')
mop_gg <- get_map(location = c(lon = mean(census$lon, na.rm = TRUE),
                               lat = mean(census$lat, na.rm = TRUE)),
                  maptype = 'satellite')

# Switch to a projected format
# Project
census_spatial <- spTransform(census_spatial, CRS( "+init=epsg:3347" ) )
mop <- spTransform(mop, CRS("+init=epsg:3347"))
# Later, we can switch back to lat/long:
census_spatial_ll <- spTransform(census_spatial, CRS("+init=epsg:4326"))
mop_ll  <- spTransform(mop, CRS("+init=epsg:4326"))
# proj4string(mop_ll) <- proj4string(zam)
```

```{r}
# Read in data at the child-specific level
children <- read_csv('redcap/CensusMopeia_DATA_2016-11-01_1734.csv')

# Keep only those households with children
children <- 
  children %>%
  filter(!is.na(perm_id_final)) %>%
  mutate(permid = perm_id_final) %>%
  dplyr::select(-perm_id_final)

# Keep only those children which appear in the non-problematic census records
children <-
  children %>%
  filter(parent_auri %in% census$uri) 

# Keep only the columns of interest
children <- 
  children %>%
  dplyr::select(permid,
                dob,
                name,
                gender,
                keeper,
                parent_auri)

# Get the household id
children <-
  children %>%
  left_join(census %>% 
              dplyr::select(house_number,
                            village_number,
                            uri),
            by = c('parent_auri' = 'uri'))
```


```{r}
if('clean_maps_with_buffers_data.RData' %in% dir()){
  load('clean_maps_with_buffers_data.RData')
} else {
  # Get spray status (will use for clusters too)
  get_spray_status() # produces spray_village (village level) and spray_status (cluster level)
  
  # Make a dataframe of just villages (with cluster assignments)
  village_df <- make_village_df(census = census,
                                spray_village = spray_village)
  
  # Get a distance matrix (takes 2-3 minutes)
  distance_matrix <- get_distance_matrix(spatial_ll_census = census_spatial_ll)
  
  # Identify problems
  identify_problems()
  census$problematic <- census$house_number %in% incorrect_houses
  census_spatial$problematic <- census_spatial$house_number %in% incorrect_houses
  census_spatial_ll$problematic <- census_spatial_ll$house_number %in% incorrect_houses
  
  # SUBSET TO REMOVE PROBLEMATIC
  # Get distance matrix after subset
  distance_matrix_old <- distance_matrix
  distance_matrix <- distance_matrix[!census$problematic,
                                         !census$problematic]
  
  # Identify the nearest neighbors of each house
  nn <- nearest_neighbor(census_spatial_ll = census_spatial_ll,
                             distance_matrix_old = distance_matrix_old,
                             n = 5)
  census$nearest_neighbors <- nn
  census_old$nearest_neighbors <- nn
  census_spatial$nearest_neighbors <- nn
  census_spatial_ll$nearest_neighbors <- nn
  
  # Remove incorrect houses
  census <- census[which(!census$house_number %in% incorrect_houses),]
  census_spatial <- census_spatial[which(!census_spatial$house_number %in% incorrect_houses),]
  census_spatial_ll <- census_spatial_ll[which(!census_spatial_ll$house_number %in% incorrect_houses),]
  
  village_df <- village_df %>% filter(!is.na(village_number),
                                      !duplicated(village_number))

  # Join spray status to objects
  joiner <- 
    village_df %>%
                dplyr::select(village_number, cluster, status)
  
  census_spatial_ll@data <-
    left_join(census_spatial_ll@data,
              joiner,
              by = 'village_number')
  census_spatial@data <-
    left_join(census_spatial@data,
              joiner,
              by = 'village_number')
  census <-
    left_join(census,
              joiner,
              by = 'village_number')
  
  # Get the voronoi tesselation surface
  vv <- voronoi(spatial_census_ll = census_spatial_ll)
  # Get the voronoi tesselation surface at the level of the spray zones
  vv_spray <- voronoi_spray_level(spatial_census_ll = census_spatial_ll)
  
  # Get a non lat long version of voronoi surface
  # For the village level
  proj4string(vv) <- proj4string(census_spatial_ll)
  vv_ll <- vv
  vv <- sp::spTransform(vv, CRSobj = proj4string(census_spatial))
  
  # For the spray-level
  proj4string(vv_spray) <- proj4string(census_spatial_ll)
  vv_spray_ll <- vv_spray
  vv_spray <- sp::spTransform(vv_spray, CRSobj = proj4string(census_spatial))
    
  # Get the buffered voronoi tesselation surface
  vvb <- rgeos::gBuffer(vv,  # village level
                        byid = TRUE, 
                        id = vv$village_number,
                        width = -1000)
  vvb_spray <- rgeos::gBuffer(vv_spray,  # village level
                        byid = TRUE, 
                        id = vv$status,
                        width = -1000)
  
  # Convert back to lat lon
  vvb_ll <- spTransform(vvb, CRSobj = proj4string(census_spatial_ll))
  vvb_ll_spray <- spTransform(vvb_spray, CRSobj = proj4string(census_spatial_ll))
  
  # Identify buffers voronoi
  census_spatial_ll <- 
    identify_buffers_voronoi(
    census_spatial_ll = census_spatial_ll,
           distance_matrix = distance_matrix,
           vvb_ll = vvb_ll)
  
  # Identify buffers voronoi at spray level
  census_spatial_ll <- 
    identify_buffers_voronoi_spray_level(census_spatial_ll = census_spatial_ll,
           distance_matrix = distance_matrix,
           vvb_ll_spray = vvb_ll_spray)
  
  # Get buffers through distance to other villages
  census_spatial_ll <- identify_buffers(spatial_ll_census = census_spatial_ll,
                       distances = distance_matrix)
  
  # Update other objects with variables only in census_spatial_ll
  update_other_objects()

  # Save
  save.image('clean_maps_with_buffers_data.RData')  
}


# Set all NA statuses to FALSE
census$status[is.na(census$status)] <- FALSE
census_spatial$status[is.na(census_spatial$status)] <- FALSE
census_spatial_ll$status[is.na(census_spatial_ll$status)] <- FALSE
village_df$status[is.na(village_df$status)] <- FALSE

# Resource helpers
source('lib/helpers.R')

# Get children's spray status
children <- 
  left_join(children,
            census %>% 
              dplyr::select(uri,
                            status),
            by = c('parent_auri' = 'uri'))
# Remove those with no spray status
children <- 
  children %>%
  filter(!is.na(status))

```


# Introduction

This document contains maps and tables pertaining to the (cleaned) Mopeia trial data. Its purpose is to explain the methods used for buffering, as well as provide visualizations and tables useful for the designation of spray / no-spray zones and study participant selection. 

It was produced the morning (GMT +2) of Thursday, November 3rd, 2016. It uses data sent from Charfudin Sacoor to Joe Brew on the evening of Sunday, October 30th, 2016 (`Casos problematticos resolvidos_Edgar Jamisse_Mopeia_2016_28_10.xlsx`).

# Methods

## Algorithmic removals

In addition to the manual cleaning carried out by the demography team, `r nrow(census_old) - nrow(census)` houses were flagged and removed in algorithmic cleaning. Houses were removed when suspected of being misclassified into the wrong village. The criteria for being of "suspicion" is somewhat complicated, but can be visualized in full here: https://github.com/joebrew/zambezia/blob/master/lib/helpers.R

## Buffering

The purpose of buffering is to flag areas which are too close to other villages to be suitable (due to the possibility of contamination). Per the protocol, there should be a 2 kilometer buffer between villages. In other words, any villager living within 1 kilometer of the _edge_ of his or her village is part of the buffer zone.

However, taking the above approach is too restrictive and it results in too great of data loss. There are many cases in which a villager may live within 1 kilometer of his or her village's edge (ie, in the "buffer"), but is still many kilometers from any other villager from another village. In these cases, there is no possibility of "contamination", so flagging that villager as in the "buffer" doesn't make sense.

In order to account for this issue, we construct delaunay triangles and subsequently create voronoi polygons from all villagers location. In short, this has the effect of "expanding" each village's boundary so that it encompasses not only those points in which the villagers live, but also any point which is closer to that village than any other.

### An example

Take village 7169, for example ('Mugurrumba'). In the below map, red points are villagers from Mugurrumba, whereas black points are residents of other villages.

```{r}
# Example
example_village <- census_spatial_ll[census_spatial_ll$village_number == 7169,]
others <- census_spatial_ll[census_spatial_ll$village_number != 7169,]

plot(example_village,
     main = 'Village 7169 (Mugurrumba)',
     col = adjustcolor('darkred', alpha.f = 0.6),
     pch = 1)
points(others, 
       col = adjustcolor('black', alpha.f = 0.5))
```

If we draw a "precise" literal border around Mugurrumba, it looks like this:

```{r}
plot(example_village,
     main = 'Village 7169 (Mugurrumba) with border',
     col = adjustcolor('darkred', alpha.f = 0.6),
     pch = 1)
points(others, 
       col = adjustcolor('black', alpha.f = 0.5), 
       add = TRUE)
x <- gConvexHull(example_village)
plot(x, 
     col = adjustcolor('darkred', alpha.f = 0.6),
     add = TRUE)
```

Note how almost all of the residents of Mugurrumba live very close (or directly on) the border

Mugurrumba is so small that it is impossible to create an _interior_ 1 kilometer buffer. In other words, _ALL_ Mugurrumba residents would be considered part of the buffer.

So, what we can do is expand Mugurrumba's "border" further out so that it includes _any_ geospatial point which is closer to Mugurrumba than any other village. This is called "voronoi tesselation", and is commonly used in geospatial applications.

```{r}
plot(example_village,
     main = 'Village 7169 (Mugurrumba) with voronoi border',
     col = adjustcolor('darkred', alpha.f = 0.6),
     pch = 1)
points(others, 
       col = adjustcolor('black', alpha.f = 0.5), 
       add = TRUE)
plot(vv_ll[vv_ll$village_number == 7169,], 
     add = T,
     col = adjustcolor('darkred', alpha.f = 0.6))
```

Now, with our "expanded" (voronoi tesselation tile) border, we can then re-draw our 1 kilometer interior border:

```{r}
plot(example_village,
     main = 'Village 7169 (Mugurrumba) with voronoi buffer',
     col = adjustcolor('darkred', alpha.f = 0.6),
     pch = 1)
points(others, 
       col = adjustcolor('black', alpha.f = 0.5))
plot(vv_ll[vv_ll$village_number == 7169,], 
     add = T,
     col = adjustcolor('black', alpha.f = 0.2),
     angle = 45,
     density = 20)
plot(vvb_ll[vvb_ll$village_number == 7169,], 
     add = T,
     col = adjustcolor('darkred', alpha.f = 0.6))
```

As can be seen in the above map, some villagers (red points outside of the red zone) are in the buffer, but many villagers (those that are not near to other villages) are not.

When we apply voronoi tesselation to the entire province of Mopeia, we _de facto_ classify every inch of territory as part of a village, even if nobody lives there. In other words, any part of Mopeia is considered part of the village which is nearest. The below map shows the entirety of the voronoi surface:

```{r}
cols <- colorRampPalette(brewer.pal(9, 'Spectral'))(nrow(vv_ll))
cols <- sample(cols, length(cols))
cols <- adjustcolor(cols, alpha.f = 0.9)
plot(mop_ll)
plot(vv_ll, col = cols,add = TRUE, border = adjustcolor('black', alpha.f = 0.5))
plot(mop_ll, add = TRUE)
```

We can apply our veronoi internal buffers and we get the following:

```{r}
cols <- colorRampPalette(brewer.pal(9, 'Spectral'))(nrow(vvb_ll))
cols <- sample(cols, length(cols))
cols <- adjustcolor(cols, alpha.f = 0.9)
plot(mop_ll)
plot(vvb_ll, col = cols,add = TRUE, border = adjustcolor('black', alpha.f = 0.5))
plot(mop_ll, add = TRUE)
```

And finally, we can add the location of each household to see which households fall into buffer zones and which don't.

```{r}
color_df <- data_frame(village_number = vvb_ll$village_number,
                       col = cols)
the_points <- census_spatial_ll 
the_points@data <- the_points@data %>%
  left_join(color_df,
            by = 'village_number')

plot(mop_ll)
plot(vvb_ll, col = adjustcolor(cols, alpha.f = 0.6),add = TRUE, border = NA)
plot(mop_ll, add = TRUE)
points(the_points,
       col = adjustcolor(the_points@data$col, alpha.f = 0.5),
       pch = 3,
       cex = 0.5)
```

Using this method, approximately half of our households are not within buffer zones. 

```{r}
ggplot(data = census_spatial_ll@data,
       aes(x = ifelse(within_1k_voronoi_buffer, 'In buffer', 'Usable'))) +
  geom_bar(alpha = 0.6) +
  xlab('Status') +
  ylab('Households') +
  ggtitle('Location status of households') +
  theme_bw()
```

### Issues with village-specific buffering

The above approach, though effective at avoiding contamination, unnecessarily delineates "borders" at the village-level, even though what we're most concerned about is contamination between spray and no-spray zones. In other words, a border and buffer between two villages of identical spray status is completely unnecessary and leads to an avoidable reduction in our number of eligible study participants.

To address this, we again use voronoi tesselation. However, instead of defining regions by village, we define only two regions: "spray" and "no-spray". Each region is multi-polygonal. Borders are defined only in the intersection of the regions, and buffers are drawn only between areas of opposite spray status.

Our spray / no-spray map looks like this:

```{r}
the_data <- vv_spray_ll
the_colors <- ifelse(the_data@data$status, 'darkred', 'darkgreen')
the_colors <- adjustcolor(the_colors, alpha.f = 0.6)
status <- the_data$status
plot(mop_ll)
plot(the_data, add = TRUE, col = the_colors, border = adjustcolor('black', alpha.f = 0.5))
legend('bottomright',
       fill = the_colors,
       legend = c('Spray', 'No spray'))
```

We add 1 kilometer internal buffers to each border to get the following:

```{r}
the_data <- vvb_ll_spray
the_colors <- ifelse(the_data@data$status, 'darkred', 'darkgreen')
the_colors <- adjustcolor(the_colors, alpha.f = 0.6)
status <- the_data$status
plot(mop_ll)
plot(the_data, col = the_colors, border = adjustcolor('black', alpha.f = 0.5), add = TRUE)
legend('bottomright',
       fill = the_colors,
       legend = c('Spray', 'No spray'))
```

We can then add each household onto our surface to visualize in which areas (spray, no-spray, or buffer) each households falls into.

```{r}
the_data <- vvb_ll_spray
the_colors <- ifelse(the_data@data$status, 'darkred', 'darkgreen')
the_colors <- adjustcolor(the_colors, alpha.f = 0.6)

plot(mop_ll)
plot(the_data, col = the_colors, border = adjustcolor('black', alpha.f = 0.5), add = TRUE)
legend('bottomright',
       fill = the_colors,
       legend = c('Spray', 'No spray'))
points(census_spatial_ll,
       col = ifelse(census_spatial_ll$within_1k_voronoi_buffer_spray, 
                    'blue',
                    ifelse(census_spatial_ll$status,
                           'darkred',
                           ifelse(!census_spatial_ll$status,
                                  'darkgreen',
                                  ifelse(is.na(census_spatial_ll$status), 'darkgreen', NA)))),
       pch = 3,
       cex = 0.3)
legend('bottomleft',
       pch = 3,
       pt.cex = 0.3,
       col = c('darkred', 'darkgreen', 'blue'),
       legend = c('Spray', 'No spray', 'Buffer'))
```



### Code 

Full code for the generation of delaunay triangles and voronoi tesselation is available at https://github.com/joebrew/zambezia.



```{r, eval = FALSE}
# VILLAGE SPECIFIC MAPS - NOT EVALUATING

maps <- lapply(village_df$row_number, function(x){
  leaflet_village_clean_with_voronoi_buffer(x)
})

htmltools::tagList(maps)
```


# Household map

The below map shows all households, except for the `r nrow(census_old) - nrow(census)` households algorithmically removed due to suspicion of error. The green areas are "no spray", the red areas are "spray" and the white areas are in buffer. 

Each point (household) is clickable; upon click, the household number, village number, core/buffer status and spray status are shown. Note that even households in white areas (buffers) received a spray / no-spray status; this can be viewed by clicking on those houses, or by viewing the "Household table" at the end of this document. 

```{r, eval = TRUE}
# leaflet_village_master_voronoi_buffer()
leaflet_village_master_voronoi_buffer_spray()
```


```{r, eval = FALSE}
# Also create a master map
master_map()
```


# Household table

The below table contains _all_ households, along with the relevant information pertaining to spray status, buffer/core status, number of children, etc. This table is interactive (ie, it can be re-ordered, etc.). Note that households not pertaining to a village with a designated spray / no-spray status are automatically marked as "no-spray".

```{r}
village_names <- 
  census %>%
  group_by(village_number, locality, local_village_name) %>%
  tally %>%
  ungroup %>%
  arrange(desc(n)) %>%
  group_by(village_number) %>%
  summarise(locality_name = first(locality),
            village_name = first(local_village_name))

# Get table for master_table, requested 25 Oct
master_table <-
  left_join(x = census_old %>% dplyr::select(-problematic),
            y = census_spatial_ll@data %>%
              filter(!duplicated(house_number)) %>%
              mutate(problematic = FALSE) %>%
              dplyr::select(house_number, problematic,
                            within_1k_voronoi_buffer_spray,
                            cluster,
                            status),
            by = 'house_number') %>%
  mutate(problematic = ifelse(is.na(problematic), 'Problematic', 'No problems suspected'),
         buffer_status = ifelse(within_1k_voronoi_buffer_spray, 'Buffer', 'Core'),
         spray_status = ifelse(status, 'SPRAY', 'NO SPRAY')) %>%
  arrange(problematic) %>%
  left_join(village_names)
master_table <- master_table[,unique(c('house_number', 
                         'village_number', 
                         'village_name',
                         'number_of_children',
                         # 'problematic', 
                         # 'nearest_neighbors',
                         'buffer_status',
                         'cluster',
                         'spray_status'))
                         # 'lng',
                         # 'lat'                         ))
                         ]
write_csv(master_table, 'household_table.csv')
names(master_table) <- toupper(gsub('_', ' ', names(master_table)))
DT::datatable(master_table)
```

# Study participant selection

There are `r nrow(children)` children with clear, correct data who are eligible for study inclusion. Of these, the split between those assigned to spray and non-spray zones is approximatey 50/50.

## Children table at village level

The number of children per village can be seen below. Note, this only includes children from households that were not flagged as "problematic" by the geospatial error detection algorithm

```{r}
child_table <- 
  children %>%
  group_by(village_number) %>%
  tally %>%
  ungroup 
child_table <- 
  left_join(child_table,
            village_names)
names(child_table) <- c('Village number',
                        'Number of children',
                        'Locality name',
                        'Village name')

DT::datatable(child_table)
write_csv(child_table, 'children_table_at_village_level.csv')
```

To assist with the randomized selection of eligible study participants, we randomize the order of the children in each village and then assign a "random_id" to each child. The "random_id" is a number between 1 and the number of children residing in that village.

```{r}
if('children_assignments.csv' %in% dir()){
  children_assignments <- read_csv('children_assignments.csv')
} else {
  set.seed(25)
  children <- children[sample(1:nrow(children), nrow(children)),]
  children_assignments <- 
    children %>%
    mutate(dummy = 1) %>%
    group_by(village_number) %>%
    mutate(random_id = cumsum(dummy)) %>%
    ungroup %>%
    dplyr::select(-dummy)
  write_csv(children_assignments,
            'children_assignments.csv')
}
children <- children_assignments
children <- children %>% arrange(village_number, random_id)

# Clean up
children <- children %>% left_join(village_names)
children <- children %>%
  dplyr::select(locality_name,
                village_name,
                village_number,
                house_number,
                random_id,
                permid,
                name,
                dob,
                gender,
                keeper,
                status)
# Get buffer status
children <-
  left_join(children,
            census_spatial_ll@data %>%
              dplyr::select(house_number,
                            buffer)) %>%
  rename(buffer_status = buffer)
```

## Children table at child level

The below table has one row for each eligible child. The `random_id` field should be used for study recruitment. Essentially, children should be recruited beginning at `random_id` 1 for each village, and going up until the minimum number of participants (18) from each village has been reached.

```{r}
DT::datatable(children)
write_csv(children, 'children_table_at_child_level.csv')
```