---
title: Cost-effectiveness evaluation of vector control strategies in Mozambique - cluster assignments
author: Joe Brew, Charfudin Sacoor, Carlos Chaccour, Molly Robertson, Francisco Saute
fig_height: 2.6
fig_width: 4
output:
  html_document:
    toc: true
    toc_float: true
    theme: yeti
---
```{r, echo = FALSE, warning = FALSE, message = FALSE, comment = NA, error= FALSE, cache = FALSE}
# No scientific notation
options(scipen=999)

# Packages 
library(xtable)
library(knitr)
library(ggplot2) 
# library(dplyr)
library(rgeos)
library(RColorBrewer)
library(Rmisc)
# Libraries
library(rgdal)
library(raster)
library(sp)
library(leaflet)
library(ggmap)
library(readxl)
library(dplyr)

# Basic knitr options
opts_chunk$set(comment = NA, 
               echo = FALSE, 
               warning = FALSE, 
               message = FALSE, 
               error = TRUE, # Render report, even with errors
               cache = F)

```

```{r setup, include=FALSE, echo = FALSE}
library(knitr)
knit_engines$set(asis = function(options) {
  if (options$echo && options$eval) knit_child(text = options$code)
})
```


```{r}
#### READ IN AND CLEAN UP DATA

# Source script
source('lib/get_zambezia.R')

# # Read in zambezia data from odk database
# get_zambezia(get_fresh = FALSE, save = FALSE)
# # Get a more nicely named core census file
# census <- CENSUS_MOPEIA_CORE

# Read in final spreadsheet, as provided by Orvalho
census <- read_excel('dropbox/final_joe_dataset.xlsx')
census <- data.frame(census)

# Get more nicely named geographic coordinate column names
census$lat <- census$y <- census$latitude <- census$gpc_lat
census$lon <- census$lng <- census$longitude <-census$x <- census$gpc_lng

# Get a shapefile for Mozambique
moz <- raster::getData('GADM', country = 'MOZ', level = 3)
moz2 <- raster::getData('GADM', country = 'MOZ', level = 2)

# Get a village number
census$village_number <- 
  as.numeric(lapply(strsplit(census$household_number, split = '-'), function(x){x[2]}))

# Create a spatial version of census
census_spatial <- census
coordinates(census_spatial) <- ~x+y
proj4string(census_spatial) <- proj4string(moz)

# Subset the shapefile just to Zambezia
zam <- moz[moz@data$NAME_1 == 'Zambezia',]

# Subset further to just Mopeia
mop <- moz2[moz2@data$NAME_2 == 'Mopeia',]

# Save the original proj4string
original_proj4string <- proj4string(moz)

# Prepare ggmap-compatible objects
zam_fortified <- fortify(zam, id = ID_3)
mop_fortified <- fortify(mop, id = ID_3)
zam_gg <- get_map(location = c(lon = mean(census$lon, na.rm = TRUE),
                               lat = mean(census$lat, na.rm = TRUE)),
                  maptype = 'satellite')
mop_gg <- get_map(location = c(lon = mean(census$lon, na.rm = TRUE),
                               lat = mean(census$lat, na.rm = TRUE)),
                  maptype = 'satellite')

# Switch to a projected format
# Project
census_spatial <- spTransform(census_spatial, CRS( "+init=epsg:3347" ) )
mop <- spTransform(mop, CRS("+init=epsg:3347"))
# Later, we can switch back to lat/long:
census_spatial_ll <- spTransform(census_spatial, CRS("+init=epsg:4326"))
mop_ll  <- spTransform(mop, CRS("+init=epsg:4326"))
# proj4string(mop_ll) <- proj4string(zam)

# Get the most prevalent village name for each village number
census_villages <-
  census %>%
  dplyr::group_by(village_number, village) %>% 
  tally %>%
  group_by(village_number) %>%
  summarise(correct_village_name = village[which(n == max(n))][1])
```

# Introduction

This document serves to explain both the methods and results for the algorithmic creation of buffered clusters for the designation of spray vs. no-spray zones for the "Cost-effectiveness evaluation of vector control strategies in Mozambique" study.

This is an internal document meant to provide an overview of methodology/approach. Its results are preliminary.

# Locations  

`r nrow(census)` locations have been identified by the census in the district of Mopeia. These locations (blue points), along with the border of the district (red), can be visualized below.

(Note: the map is interactive. Zoom in for a closer look, or click on a blue dot to see the house ID number and village name.)


```{r}
leaflet(mop_ll) %>%
  addProviderTiles("Esri.WorldImagery") %>%
  addPolylines(color = 'red') %>%
  # addPolygons(stroke = FALSE, 
  #             smoothFactor = 0.2, 
  #             fillOpacity = 0.25,
  #             col = 'red') %>%
  addCircleMarkers(lng = census$lng,
  lat = census$lat,
  color = colors,
  fillColor = 'blue',
  radius = 2.5,
  opacity = 0,
  fillOpacity = 0.5,
  popup = paste0('Household: ', census$household_number, ' Village number: ',
  census$village_number))
```


# Parameters

We must designate approximately 86 clusters for random assignation to "spray" / "no-spray" groups. These clusters have no size restrictions, but are subject to multiple constraints:

- Assignation must not divide villages (ie, one village must have exactly 1 spray/no-spray status).
- Each cluster must contain a buffer of at least 1 kilometer.
- Buffers may not overlap (effectively making the buffer size 2 kilometers).
- Units near the rim of the distrct should be excluded (for logistical purposes).
- There must be at least 18 children <5 years of age in each core area (at time of enrollment).
- Those children should be from different houses, ideally.
- The 86 clusters must be divided into 3 strata (with similarly sized clusters per strata).


# Incorrectly classified households  

## Identifying errors

It appears that many households are not assigned the right village number, given their geographic location. To detect these, we take the distance from each household to all the other households in its village; in the case that the median distance from the household to all the other households exceeds 20,000 UTM degrees, it is considered "suspected" to be incorrect. 

The below table shows all 191 villages. The `n_incorrect_houses` column indicates the number of households in each village which are suspected to _not_ be part of that village.


```{r}
# Get the area and other metrics for each village
unique_village_numbers <- sort(unique(census$village_number))
village_df <- census %>%
  group_by(village_number) %>%
  summarise(n_children = sum(number_of_children),
            houses_with_any_children = length(which(number_of_children > 0)),
            houses = n(),
            area = NA,
            incorrect_houses = NA,
            n_incorrect_houses = NA)

# Create placeholder vector of incorrect houses
incorrect_houses <- c()
  
color <- 'darkgreen'
for (i in 1:nrow(village_df)){
  message(i)
  # Get village name
  this_village <- village_df$village_number[i]
    # Get the points only
  sub_census <- 
    census_spatial[census_spatial$village_number == this_village,]
  # Get the border
  border <- gConvexHull(sub_census)
  # Calculate area
  this_area <- gArea(border)
  # Populate area in dataframe
  village_df$area[i] <- this_area
  
  # Go through each household in the unit and calculate distance
  # to all other units
  distances <- gDistance(sub_census, byid = TRUE)
  median_distances <- apply(distances, 1, median)

  # From the vector of median distances, flag those which are abnormal
  threshold <- 20000
  incorrect_house_ids <- sub_census$household_number[median_distances > threshold]
  # Stick results into dataframe
  village_df$n_incorrect_houses[i] <- length(incorrect_house_ids)
  village_df$incorrect_houses[i] <- paste0(incorrect_house_ids, collapse = ', ')
  
  # Add to the vector of incorrect houses 
  incorrect_houses <- c(incorrect_houses, incorrect_house_ids)
}

# Reorder the village dataframe by size
village_df <- 
  village_df %>%
  arrange(desc(houses))

# Print table
x <- village_df %>%
  dplyr::select(-incorrect_houses)
DT::datatable(x)
```

All in all, `r length(incorrect_houses)` houses are suspected to be incorrectly classified (ie, either their geographic coordinates are incorrect, or their household ID number is incorrect). These are those IDs:

```{r}
incorrect_houses_df <- data.frame(household_number = incorrect_houses)
incorrect_houses_df <-
  incorrect_houses_df %>%
  left_join(census %>%
              dplyr::select(household_number, village_number)) %>%
  arrange(village_number) %>%
  dplyr::select(village_number, household_number)
DT::datatable(incorrect_houses_df)
```

All in all, `r length(which(village_df$incorrect_houses > 0))` of `r nrow(village_df)` villages appear to have incorrect household inclusions. The below maps show each of these `r length(which(village_df$incorrect_houses > 0))` villages. Note the abnormal (and large) polygon shapes, due to the incorrect inclusions.

```{r}
incorrect_villages <- village_df[which(village_df$incorrect_houses > 0),]
for(i in 1:nrow(incorrect_villages)){
  this_village_number <- incorrect_villages$village_number[i]
   # Get the points only
  sub_census <- census_spatial_ll[census_spatial_ll$village_number == this_village_number,]
  # Plot Mopeia
  plot(mop_ll)
  # Get the border
  border <- gConvexHull(sub_census)
  # Plot the border
  plot(border, add = TRUE, col = adjustcolor(color, alpha.f = 0.3),
       border = adjustcolor(color, alpha.f = 0.8))
  # Add the points
  points(sub_census,
         pch = '.',
         col = adjustcolor(color, alpha.f = 0.6))
  # Add title
  title(main = paste0('Village number: ', unique_village_numbers[i]))
}
```

## Addressing errors

In order to address errors, we simply remove the `r length(incorrect_houses)` houses we suspect to be incorrectly classified.  

```{r}
census <- census %>%
  filter(!household_number %in% incorrect_houses)

# Create a spatial version of census
census_spatial <- census
coordinates(census_spatial) <- ~x+y
proj4string(census_spatial) <- proj4string(moz)
# Switch to a projected format
# Project
census_spatial <- spTransform(census_spatial, CRS( "+init=epsg:3347" ) )
# Later, we can switch back to lat/long:
census_spatial_ll <- spTransform(census_spatial, CRS("+init=epsg:4326"))

```


## Combining into clusters

This leaves `r nrow(census)` households from a total of `r length(unique(census$village_number))` villages. 

Having cleaned our data, we now want to combine any villages which are _too_ small (ie, have fewer than 18 children). 

There are `r length(which(village_df$n_children < 18))` of these. In order to account for them, we simply add them to the most nearby village which _does_ have enough children.

```{r}
# Get centroids
village_df <-
  village_df %>%
  left_join(census %>%
  group_by(village_number) %>%
  summarise(x_centroid = mean(lng),
            y_centroid = mean(lat)))

# Get distances between villages
distances <- spDists(x = as.matrix(village_df %>% dplyr::select(x_centroid, y_centroid)),
        longlat = TRUE,
        diagonal = FALSE)

# Assign preliminary clusters
counter <- 0
village_df$cluster <- NA
for (i in 1:nrow(village_df)){
  if(village_df$n_children[i] >= 18){
    counter <- counter + 1
    village_df$cluster[i] <- counter
  }
}

# Go through each row, and if not enough children, re-assign to nearby one with enough children
enough_children <- village_df$n_children >= 18
for (i in 1:nrow(village_df)){
  # If not enough children
  if(is.na(village_df$cluster[i])){
    # Get the nearest village which isn't too small
    possibles <- data.frame(cluster = village_df$cluster,
                            distance = distances[i,])
    possibles <- possibles %>%
      filter(!is.na(cluster))
    this_cluster <- possibles$cluster[which.min(possibles$distance)]
    village_df$cluster[i] <- this_cluster
  }
}

# Bring cluster type into census
census <- census %>% 
  left_join(village_df %>%
              dplyr::select(village_number, cluster),
            by = 'village_number')
# Create a spatial version of census
census_spatial <- census
coordinates(census_spatial) <- ~x+y
proj4string(census_spatial) <- proj4string(moz)
# Switch to a projected format
# Project
census_spatial <- spTransform(census_spatial, CRS( "+init=epsg:3347" ) )
# Later, we can switch back to lat/long:
census_spatial_ll <- spTransform(census_spatial, CRS("+init=epsg:4326"))
```

This leaves us with `r length(unique(village_df$cluster))` clusters.  These clusters can be visualized below.

```{r}
plot(mop_ll)

possible_colors <- rainbow(1000)
clusters <- unique(sort(village_df$cluster))
for (i in 1:length(clusters)){
  this_cluster <- clusters[i]
  these_villages <- village_df$village_number[village_df$cluster == this_cluster]
  these_houses <- census_spatial_ll[census_spatial_ll$village_number %in% these_villages,]
  # Get Convex hull
  border <- gConvexHull(these_houses)
  # Get a random color
  the_color <- sample(possible_colors, 1)
  plot(border, add = T,
       col = adjustcolor(the_color, alpha.f = 0.1),
       border = adjustcolor(the_color, alpha.f = 0.5))
  points(these_houses, pch = '.',
         col = adjustcolor(the_color, alpha.f = 0.8))
}
```

# Cluster assignation

The breakdown of cluster assignations can be viewed below.

```{r}
x <- village_df %>%
  group_by(cluster) %>%
  summarise(village_numbers = paste0(village_number, collapse = ', '))
DT::datatable(x)
```

# Strata assignation  

We now want to divide our clusters into three kinds: small, medium, and large. For this, we use simple terciles to bin the number of houses per cluster.

```{r}
cluster_df <- census %>%
  group_by(cluster) %>%
  summarise(n_children = sum(number_of_children),
            houses_with_any_children = length(which(number_of_children > 0)),
            houses = n())

# Bin the houses
the_breaks <- quantile(cluster_df$houses, probs = c(0, 0.33, 0.66, 1))
bins <- cut(cluster_df$houses, 
            breaks = the_breaks,
            include.lowest = TRUE,
            labels = c('Small', 'Medium', 'Large'))
cluster_df$strata <- bins
```

The strata category cut-offs are 0-`r round(as.numeric(the_breaks[2]))` (small), `r as.numeric(the_breaks[2])`-`r as.numeric(the_breaks[3])` (medium) and `r as.numeric(the_breaks[2])` or more (large). These cut-offs can be visualized in the below histogram.

```{r}
hist(cluster_df$houses,
     breaks = 100,
     main = 'Distribution of number of households per cluster',
     border = NA,
     col = 'grey',
     xlab = 'Houses')
abline(v = as.numeric(the_breaks[2:3]),
       col = adjustcolor('darkred', alpha.f = 0.6))
```

The strata for each village is viewable in the below table

```{r}
DT::datatable(cluster_df)
```

# Spray / no-spray assignation

From the `r nrow(cluster_df)` clusters, we randomly assign all to spray or no-spray status, using our size strata so that an equal number of small, medium, and large clusters get each status. The assignation is as follows.

```{r}
set.seed(1000)
results_list <- list()
stratas <- c('Small', 'Medium', 'Large')
for (i in 1:length(stratas)){
  this_strata <- cluster_df %>% 
    filter(strata == stratas[i])
  # Get exactly half to certain status
  n_assign <- round(0.5 * nrow(this_strata))
  # Randomly get n_assign rows
  these_ones <- sample(1:nrow(this_strata), n_assign)
  this_strata$status <- 'No spray'
  this_strata$status[these_ones] <- 'Spray'
  results_list[[i]] <- this_strata
}
cluster_df <- do.call('rbind', results_list)
```

The below table shows the spray / no-spray status _by cluster_.

```{r}
x <- cluster_df
DT::datatable(x)
```

The below table shows the spray / no-spray status _by village_.

```{r}
x <- 
  census %>%
  group_by(village_number) %>%
  summarise(houses = n(),
            cluster = first(cluster)) %>%
  left_join(cluster_df %>%
              dplyr::select(cluster, status))
DT::datatable(x)
# Write a csv
readr::write_csv(x, '~/Desktop/spray_no_spray_assignments.csv')
```

The spray / no-spray statuses can be visualized in the below maps. Blue is no spray; red is spray.

```{r}
# Bring the statuses into census
census <- census %>% 
  left_join(cluster_df %>%
              dplyr::select(cluster, status, strata),
            by = 'cluster')
# Create a spatial version of census
census_spatial <- census
coordinates(census_spatial) <- ~x+y
proj4string(census_spatial) <- proj4string(moz)
# Switch to a projected format
# Project
census_spatial <- spTransform(census_spatial, CRS( "+init=epsg:3347" ) )
# Later, we can switch back to lat/long:
census_spatial_ll <- spTransform(census_spatial, CRS("+init=epsg:4326"))

# # PLOT IT
# plot(mop_ll)
# cols <- ifelse(census$status == 'Spray', 'darkgreen', 'darkred')
# points(census_spatial_ll,
#        col = cols,
#        pch = '.')

# Divide into spray / no-spray
spray <- census %>% filter(status == 'Spray')
no_spray <- census %>% filter(status == 'No spray')

leaflet() %>%
  addProviderTiles("Esri.WorldImagery") %>%
  addPolylines(data = mop_ll, color = 'black') %>%
  addCircleMarkers(lng = spray$lng,
  lat = spray$lat,
  color = colors,
  fillColor = 'red',
  radius = 2.5,
  opacity = 0,
  fillOpacity = 0.5,
  popup = paste0('Household: ', spray$household_number, ' Village number: ',
  spray$village_number, ' Cluster number: ', spray$cluster)) %>%
  
  addCircleMarkers(lng = no_spray$lng,
  lat = no_spray$lat,
  color = colors,
  fillColor = 'blue',
  radius = 2.5,
  opacity = 0,
  fillOpacity = 0.5,
popup = paste0('Household: ', no_spray$household_number, ' Village number: ',
  no_spray$village_number, ' Cluster number: ', no_spray$cluster)) 
  
```


# Caveats

1. This approach uses some "automated" data cleaning. It has not been validated.

2. This approach yields `r (length(unique(sort(x$cluster))))` clusters, which is far more than the required. The assumption herein is that some manual "elimination" of clusters will allow for the satisfication of some operational parameters (ie, avoiding Renamo-occupied villages, villages that are difficult to reach, etc.). 

