---
title: "Cluster selection: distance to health posts and other issues of concern"
author: "Joe Brew"
fig_width: 4
output:
  html_document:
    theme: flatly
    toc: yes
    toc_float: yes
  pdf_document:
    toc: yes
fig_height: 2.6
---
```{r, echo = FALSE, warning = FALSE, message = FALSE, comment = NA, error= FALSE, cache = FALSE}
# No scientific notation
options(scipen=999)

# Packages 
library(xtable)
library(knitr)
library(ggplot2) 
# library(dplyr)
library(rgeos)
library(RColorBrewer)
library(Rmisc)
# Libraries
library(rgdal)
library(raster)
library(sp)
library(leaflet)
library(ggmap)
library(readxl)
library(dplyr)
library(readr)
library(geosphere)

# Basic knitr options
opts_chunk$set(comment = NA, 
               echo = FALSE, 
               warning = FALSE, 
               message = FALSE, 
               error = TRUE, # Render report, even with errors
               cache = F)

# Helpers
source('lib/get_zambezia.R')
source('lib/helpers.R')
```

```{r setup, include=FALSE, echo = FALSE}
library(knitr)
knit_engines$set(asis = function(options) {
  if (options$echo && options$eval) knit_child(text = options$code)
})
```


```{r, message=FALSE}
#### READ IN AND CLEAN UP DATA

# Read in final dataset, as given by Orvahlo on October 24, 2016
# census <- read_excel('Limpesa_Final__241016_For maps.xlsx')
# Read in corrected final dataset, sent by Edgar on October 30, 2016
# census <- read_csv('Casos problematticos resolvidos_Edgar Jamisse_Mopeia_2016_28_10.csv')

census <- read_csv('redcap/CensusMopeia_DATA_2016-11-01_1734.csv')

# Divide between children and census
census$child <- census$redcap_event_name == 'census_mopeia_chil_arm_1'
children <- census %>% filter(child)
census <-
  census %>% filter(!child)

# Fix weird header
names(census)[6] <- 'locality_Final'

# Get more nicely named geographic coordinate column names
census$lat <- census$y <- census$latitude <- census$gpc_lat
census$lon <- census$lng <- census$longitude <-census$x <- census$gpc_lng

# Get a shapefile for Mozambique
moz <- raster::getData('GADM', country = 'MOZ', level = 3)
moz2 <- raster::getData('GADM', country = 'MOZ', level = 2)

# GET CORRECT FINAL COLUMNS, AS EXPLAINED IN EMAIL
census$house_number <- census$houseno_final
census$permid <- census$perm_id_final
census$village_number <-
  unlist(lapply(strsplit(census$house_number, split = '-'), function(x){x[2]}))

# Convert village number to numeric
census$village_number <- 
  as.numeric(as.character(census$village_number))

# Keep only necessary vars in children
children <- 
  children %>%
  dplyr::select(parent_auri,
                permid,
                name:children_info_complete)

# Keep only necessary info in census
census <- census %>%
  dplyr::select(uri:family_id,
                latitude:village_number)

# Clean up geography
census$x <- census$lng <- census$gpc_lng
census$y <- census$lat <- census$gpc_lat

# Remove all NA rows
census <- census %>%
  filter(!is.na(lng),
         !is.na(lat),
         !is.na(village_number),
         !is.na(house_number))


# Save old version of census
census_old <- census

# Create a spatial version of census
census_spatial <- census
coordinates(census_spatial) <- ~x+y
proj4string(census_spatial) <- proj4string(moz)

# Subset the shapefile just to Zambezia
zam <- moz[moz@data$NAME_1 == 'Zambezia',]

# Subset further to just Mopeia
mop <- moz2[moz2@data$NAME_2 == 'Mopeia',]

# Save the original proj4string
original_proj4string <- proj4string(moz)

# Prepare ggmap-compatible objects
zam_fortified <- fortify(zam, id = ID_3)
mop_fortified <- fortify(mop, id = ID_3)
zam_gg <- get_map(location = c(lon = mean(census$lon, na.rm = TRUE),
                               lat = mean(census$lat, na.rm = TRUE)),
                  maptype = 'satellite')
mop_gg <- get_map(location = c(lon = mean(census$lon, na.rm = TRUE),
                               lat = mean(census$lat, na.rm = TRUE)),
                  maptype = 'satellite')

# Switch to a projected format
# Project
census_spatial <- spTransform(census_spatial, CRS( "+init=epsg:3347" ) )
mop <- spTransform(mop, CRS("+init=epsg:3347"))
# Later, we can switch back to lat/long:
census_spatial_ll <- spTransform(census_spatial, CRS("+init=epsg:4326"))
mop_ll  <- spTransform(mop, CRS("+init=epsg:4326"))
# proj4string(mop_ll) <- proj4string(zam)
```

```{r}
# Read in data at the child-specific level
# children <- read_csv('redcap/CensusMopeia_DATA_2016-11-01_1734.csv')

# Keep only those children which appear in the non-problematic census records
children <-
  children %>%
  filter(parent_auri %in% census$uri) 

# Keep only the columns of interest
children <- 
  children %>%
  dplyr::select(permid,
                dob,
                name,
                gender,
                keeper,
                parent_auri)

# Get the household id
children <-
  children %>%
  left_join(census %>% 
              dplyr::select(house_number,
                            village_number,
                            uri),
            by = c('parent_auri' = 'uri'))
```


```{r}
if('clean_maps_with_buffers_data.RData' %in% dir()){
  load('clean_maps_with_buffers_data.RData')
} else {
  # Get spray status (will use for clusters too)
  get_spray_status() # produces spray_village (village level) and spray_status (cluster level)
  
  # Make a dataframe of just villages (with cluster assignments)
  village_df <- make_village_df(census = census,
                                spray_village = spray_village)
  
  # Get a distance matrix (takes 2-3 minutes)
  distance_matrix <- get_distance_matrix(spatial_ll_census = census_spatial_ll)
  
  # Identify problems
  identify_problems()
  census$problematic <- census$house_number %in% incorrect_houses
  census_spatial$problematic <- census_spatial$house_number %in% incorrect_houses
  census_spatial_ll$problematic <- census_spatial_ll$house_number %in% incorrect_houses
  
  # SUBSET TO REMOVE PROBLEMATIC
  # Get distance matrix after subset
  distance_matrix_old <- distance_matrix
  distance_matrix <- distance_matrix[!census$problematic,
                                         !census$problematic]
  
  # Identify the nearest neighbors of each house
  nn <- nearest_neighbor(census_spatial_ll = census_spatial_ll,
                             distance_matrix_old = distance_matrix_old,
                             n = 5)
  census$nearest_neighbors <- nn
  census_old$nearest_neighbors <- nn
  census_spatial$nearest_neighbors <- nn
  census_spatial_ll$nearest_neighbors <- nn
  
  # Remove incorrect houses
  census <- census[which(!census$house_number %in% incorrect_houses),]
  census_spatial <- census_spatial[which(!census_spatial$house_number %in% incorrect_houses),]
  census_spatial_ll <- census_spatial_ll[which(!census_spatial_ll$house_number %in% incorrect_houses),]
  
  village_df <- village_df %>% filter(!is.na(village_number),
                                      !duplicated(village_number))

  # Join spray status to objects
  joiner <- 
    village_df %>%
                dplyr::select(village_number, cluster, status)
  
  census_spatial_ll@data <-
    left_join(census_spatial_ll@data,
              joiner,
              by = 'village_number')
  census_spatial@data <-
    left_join(census_spatial@data,
              joiner,
              by = 'village_number')
  census <-
    left_join(census,
              joiner,
              by = 'village_number')
  
  # Get the voronoi tesselation surface
  vv <- voronoi(spatial_census_ll = census_spatial_ll)
  # Get the voronoi tesselation surface at the level of the spray zones
  vv_spray <- voronoi_spray_level(spatial_census_ll = census_spatial_ll)
  
  # Get a non lat long version of voronoi surface
  # For the village level
  proj4string(vv) <- proj4string(census_spatial_ll)
  vv_ll <- vv
  vv <- sp::spTransform(vv, CRSobj = proj4string(census_spatial))
  
  # For the spray-level
  proj4string(vv_spray) <- proj4string(census_spatial_ll)
  vv_spray_ll <- vv_spray
  vv_spray <- sp::spTransform(vv_spray, CRSobj = proj4string(census_spatial))
    
  # Get the buffered voronoi tesselation surface
  vvb <- rgeos::gBuffer(vv,  # village level
                        byid = TRUE, 
                        id = vv$village_number,
                        width = -1000)
  vvb_spray <- rgeos::gBuffer(vv_spray,  # village level
                        byid = TRUE, 
                        id = vv$status,
                        width = -1000)
  
  # Convert back to lat lon
  vvb_ll <- spTransform(vvb, CRSobj = proj4string(census_spatial_ll))
  vvb_ll_spray <- spTransform(vvb_spray, CRSobj = proj4string(census_spatial_ll))
  
  # Identify buffers voronoi
  census_spatial_ll <- 
    identify_buffers_voronoi(
    census_spatial_ll = census_spatial_ll,
           distance_matrix = distance_matrix,
           vvb_ll = vvb_ll)
  
  # Identify buffers voronoi at spray level
  census_spatial_ll <- 
    identify_buffers_voronoi_spray_level(census_spatial_ll = census_spatial_ll,
           distance_matrix = distance_matrix,
           vvb_ll_spray = vvb_ll_spray)
  
  # Get buffers through distance to other villages
  census_spatial_ll <- identify_buffers(spatial_ll_census = census_spatial_ll,
                       distances = distance_matrix)
  
  # Update other objects with variables only in census_spatial_ll
  update_other_objects()

  # Save
  save.image('clean_maps_with_buffers_data.RData')  
}


# Set all NA statuses to FALSE
census$status[is.na(census$status)] <- FALSE
census_spatial$status[is.na(census_spatial$status)] <- FALSE
census_spatial_ll$status[is.na(census_spatial_ll$status)] <- FALSE
village_df$status[is.na(village_df$status)] <- FALSE

# Resource helpers
source('lib/helpers.R')

# Get children's spray status
children <- 
  left_join(children,
            census %>% 
              dplyr::select(uri,
                            status),
            by = c('parent_auri' = 'uri'))
# Remove those with no spray status
children <- 
  children %>%
  filter(!is.na(status))
```

```{r}
# Read in the coordinates of health centers
us <- read.csv('unidades_sanitarias_cleaned.csv')
# Make spatial
us$x <- us$lng <- us$lon <- us$longitude <- us$gpc_lng
us$y <- us$lat <- us$latitude <- us$gpc_lat
coordinates(us) <- ~gpc_lng+gpc_lat
```

# Health facility locations

## Table

The below table shows the coordinates, localidades and postos administrativos of the 12 health facilities in Mopeia.

```{r}
x <- us@data
x <- x %>% dplyr::select(health_facility,
                         loc_health_facility,
                         pos_adm_health_facility,
                         lng,
                         lat)
x$lng <- round(x$lng, 4)
x$lat <- round(x$lat, 4)
names(x) <- c('Facility', 'Localidade', 'Posto Adm.', 'Lng', 'Lat')
DT::datatable(x)
rm(x)
```

## Map

The below map shows the locations of each health facility. Clicking on the blue markers reveals the name of each health facility.

```{r}

leaflet(data = us@data) %>%
  addProviderTiles('OpenStreetMap.BlackAndWhite') %>%
  addMarkers(~lng,
             ~lat,
             popup = ~health_facility) %>%
  addPolygons(data = mop_ll, stroke = FALSE, color = 'red') 

```

# Cluster locations

```{r}
# Get clusters
clusters <-
  census_spatial_ll@data %>%
  group_by(cluster) %>%
  summarise(lng = mean(lng),
            lat = mean(lat),
            spray_status = first(status),
            houses = n(),
            houses_core = length(which(!within_1k_voronoi_buffer_spray)),
            houses_buffer = length(which(within_1k_voronoi_buffer_spray)),
            children = sum(number_of_children),
            children_core = sum(number_of_children[!within_1k_voronoi_buffer_spray]),
            children_buffer = sum(number_of_children[within_1k_voronoi_buffer_spray]),
            villages = length(unique(village_number)),
            village_numbers = paste0(sort(unique(village_number)), collapse = ', ')) %>%
  filter(!is.na(cluster)) %>%
  mutate(spray_status = ifelse(spray_status, 'Spray', 'No spray'))
```

There were originally `r nrow(clusters)` clusters.

```{r}
# filter out < 22
clusters <- clusters %>%
  filter(children_core >= 22)
```
We filter out any cluster with fewer than 22 children, and are left with `r nrow(clusters)` clusters.



## Table

The below table shows the centroid of each cluster, along with some descriptive information regarding the inhabitants.

```{r}
DT::datatable(clusters %>%
                mutate(lng = round(lng, 3),
                       lat = round(lat, 3)))
```

## Map

The below map shows each cluster's centroid (circles) along with each health facility (pins). Both pins and circles reveal their identity when clicked.

```{r}
leaflet() %>%
  addProviderTiles('OpenStreetMap.BlackAndWhite') %>%   
  addPolygons(data = mop_ll, stroke = FALSE, color = 'red') %>%
  addMarkers(data = us@data,
             ~lng,
             ~lat,
             popup = ~health_facility) %>%
  addCircleMarkers(data = clusters,
                   lng = clusters$lng,
                   lat = clusters$lat,
                   color = 'darkgreen',
                   fillOpacity = 0.8,
                   radius = 3,
                   opacity = 0,
                   popup = as.character(clusters$cluster))
```


## Distance calculations

First, we need to calculate the distance in kilometers between each clusters' centroid and each health post, in order to get the nearest health post. We do this using Levenstein "great-circle" distance. Note, this does not take into account travel times, roads, or geographical barriers such as rivers.

```{r}
clusters_sp <- clusters
clusters_sp$x <- clusters_sp$lng
clusters_sp$y <- clusters_sp$lat
coordinates(clusters_sp) <- ~x+y
proj4string(clusters_sp) <- proj4string(us) <- proj4string(census_spatial_ll)
us_distances <- geosphere::distm(x = us,
                              y = clusters_sp, 
                              fun = distVincentySphere)
# Get which is the closest
clusters$closest_health_facility <-
  apply(us_distances, 2, function(x){
    us$health_facility[which.min(x)]
  })
# Get distance
clusters$km_to_closest_health_facility <-
  apply(us_distances, 2, function(x){
    min(x) / 1000
  })

# FOR US, get distance from noere
distance_from_noere <- 
  geosphere::distm(x = us,
                   y = us[grepl('NOERE', us@data$health_facility),])
us$km_from_noere <- as.numeric(distance_from_noere / 1000)
# Arrange by distance from noere, so that we appropriately keep clusters together
us <- us[order(us@data$km_from_noere),]
```

What follows is a table of each cluster, its nearest health facility, and the distance in kilometers between that health facility and the cluster centroid.

```{r}
DT::datatable(clusters %>%
                dplyr::select(cluster,
                              closest_health_facility,
                              km_to_closest_health_facility) %>%
                arrange(desc(km_to_closest_health_facility)))
```

Distance between cluster centroids and their respective nearest health posts vary widely. The below histogram shows the distribution in distance (kilometers) to each centroids' nearest health post.

```{r}
ggplot(data = clusters,
       aes(x = km_to_closest_health_facility)) +
  geom_histogram(alpha = 0.6, fill = 'darkgreen') +
  theme_get() +
  xlab('Kilometers') +
  ylab('Clusters') +
  ggtitle('Distance from cluster centroid to nearest health post')
nearest_97 <- as.numeric(round(quantile(clusters$km_to_closest_health_facility, 0.97), digits = 2))
```

97% of all clusters are within `r nearest_97` kilometers. Therefore, for logstical purposes, we will eliminate those 3% which are beyond that distance.

```{r}
clusters <- clusters %>%
  mutate(remove = km_to_closest_health_facility > nearest_97)

clusters_sp@data <- left_join(clusters_sp@data,
                              clusters)
```

Following elimination, our remaining clusters and health posts look like this. Points in red are clusters which are not eligible for selection (ie, they are removed), due to the far distance from a health post.

```{r}
remove_these <- clusters %>% filter(remove)
keep_these <- clusters %>% filter(!remove)
leaflet() %>%
  addProviderTiles('OpenStreetMap.BlackAndWhite') %>%   
  addPolygons(data = mop_ll, stroke = FALSE, color = 'red') %>%
  addMarkers(data = us@data,
             ~lng,
             ~lat,
             popup = ~health_facility) %>%
  addCircleMarkers(data = remove_these,
                   lng = remove_these$lng,
                   lat = remove_these$lat,
                   color = 'darkred',
                   fillOpacity = 0.8,
                   radius = 3,
                   opacity = 0,
                   popup = as.character(remove_these$cluster)) %>%
  addCircleMarkers(data = keep_these,
                   lng = keep_these$lng,
                   lat = keep_these$lat,
                   color = 'darkgreen',
                   fillOpacity = 0.8,
                   radius = 3,
                   opacity = 0,
                   popup = as.character(keep_these$cluster))
```

```{r}
# Make the removals
clusters <- clusters %>%
  filter(!remove)

clusters_sp <- clusters_sp[which(!clusters_sp$remove),]
```

Following elimination based on distance, the number of clusters eligible for selection goes down to `r nrow(clusters)`.

# Random selection

## Strata size

Having eliminated the 25% of clusters which are furthest from their respective health posts, we carry out a random selection. We use terciles of number of houses to divide our remaining clusters into three sizes.

```{r}
cut_offs <- quantile(clusters$houses, probs = c(0, .33, .66, 1))
strata <- base::cut(clusters$houses, breaks = c(0, cut_offs[2:3], Inf))
clusters$strata <- strata
clusters$size <- c('Small', 'Medium', 'Large')[as.numeric(clusters$strata)]
```

The cut-off points for each strata are:

- Small: 0 to `r as.numeric(floor(cut_offs[2]))` houses.
- Medium: `r as.numeric(ceiling(cut_offs[2]))` to `r as.numeric(floor(cut_offs[3]))` houses.
- Large: `r as.numeric(floor(cut_offs[3]))` or more houses.

We can visualize the distribution of clusters' numbers of houses below, along with lines indicating the cut points:

```{r}
ggplot(data = clusters,
       aes(x = houses)) +
  geom_density(aes(fill = size),
    # fill = 'darkorange', 
               alpha = 0.6, color = NA) +
  geom_vline(xintercept = as.numeric(cut_offs),
             color = 'darkred',
             alpha = 0.7, 
             lty = 3) +
  xlab('Houses') +
  ylab('Density') +
  ggtitle('Distribution of number of houses',
          'Cluster level') +
  scale_fill_manual(name = 'Size',
                    values = c('darkblue', 'red', 'darkorange'))
```

## Selection

Ultimately, we need to select 86 clusters, stratified by size, and health post, and in equal number (43) between spray and no spray status. In other words, we want to select the below number of clusters:

```{r}
selector <- expand.grid(spray_status = c('Spray', 'No spray'),
                        size = c('Small', 'Medium', 'Large'))
selector$clusters <- c(rep(14, 4), rep(15, 2))
selector
```

Cluster assignment (and working groups) hinge on health facilities. Prior to selection, using the `r nrow(clusters)` which are eligible for selection, the number of clusters per health facility is as follows:

```{r}
x <- clusters %>%
  group_by(health_facility = closest_health_facility) %>%
  summarise(n_clusters = n(),
            clusters = paste0(sort(cluster), collapse = ', ')) %>%
  arrange(n_clusters)
```

As is apparent above, some health facilities have fewer than 5 associated eligible clusters. 

We do an initial pass to ensure that all health facilities get at least 5 clusters (when possible) or if there are not 5 clusters belonging to that health facility, the number of clusters which are closest to that health facility (as low as 1). In a second pass, we randomly pick from the remaining clusters stratified by size and spray status.

We do this randomly and get the following selections:

```{r}
set.seed(123)
# re arrange
clusters <- clusters[sample(1:nrow(clusters), nrow(clusters)),]
# randomly select
clusters$selected <- FALSE
# Assign as "selected" any cluster belonging to a health
# facility with five or fewer clusters
clusters <- 
  clusters %>%
  group_by(closest_health_facility) %>%
  mutate(dummy = n()) %>%
  ungroup %>%
  mutate(selected = ifelse(dummy <= 5, TRUE, selected))

# Since we've made an inital prior selection, add to the counters
adjust_counter <-
  clusters %>% 
  filter(selected) %>% 
  group_by(size, spray_status) %>% 
  summarise(counter = n())
# Join the adjustments to selector
selector <- left_join(selector, adjust_counter)
selector$counter[is.na(selector$counter)] <- 0

for (i in 1:nrow(clusters)){
  # Get the cluster in question
  this_cluster <- clusters[i,]
  # Ensure that the number isn't yet above the max
  if(!this_cluster$selected){
    this_selector <- selector %>% 
    filter(spray_status == this_cluster$spray_status,
           size == this_cluster$size)
  if(this_selector$counter < this_selector$clusters){
  clusters$selected[i] <- TRUE
  selector$counter[selector$spray_status == this_cluster$spray_status &
                   selector$size == this_cluster$size] <- 
    this_selector$counter +1
  }
}
}
```

```{r}
cat(paste0(sort(unique(clusters$cluster[clusters$selected])), collapse = ', '))
```

Below is a table of all clusters, including both those selected and those not selected. The `health_facility` variable refers to the health facility which is nearest to that cluster. The `km` variable is the distance in (Levenstein) kilometers from the cluster's centroid to that health facility.

```{r}
DT::datatable(clusters %>%
                dplyr::select(closest_health_facility,
                              cluster,
                              spray_status,
                              size,
                              selected, 
                              km_to_closest_health_facility) %>%
  arrange(desc(selected, size, spray_status)) %>%
  arrange(desc(closest_health_facility)) %>%
  left_join(children %>%
              left_join(census_spatial_ll@data %>%
                        dplyr::select(uri,
                                      within_1k_voronoi_buffer_spray,
                                      cluster) %>%
                          filter(!within_1k_voronoi_buffer_spray),
                        by = c('parent_auri' = 'uri')) %>%
              group_by(cluster) %>%
              summarise(n_children_in_core = n())) %>%
    mutate(km_to_closest_health_facility = round(km_to_closest_health_facility, digits = 1))
  ) 
```

# Assignation to health posts and workers

Clusters will be grouped and assigned to workers. Where each worker will be based will the health post nearest to that workers' assigned clusters.


## Segregation into sub-groups

Having randomly selected clusters, we can see how many clusters are assigned to each health facility. The previous concept of "groups" has been abandoned, since the random assignments are not divisable by group size (4/5) in many cases.

```{r}
x = clusters %>%
  filter(selected) %>%
  group_by(health_facility = closest_health_facility) %>%
  summarise(n_clusters = n(),
            clusters = paste0(sort(cluster), collapse = ', ')) %>%
  arrange(desc(n_clusters)) 
  
DT::datatable(x)
# ordered_clusters <- us@data %>%
#   dplyr::select(health_facility) %>%
#   left_join(clusters %>%
#               filter(selected) %>%
#               group_by(health_facility = closest_health_facility) %>%
#               mutate(dummy = 1)) %>%
#   filter(!is.na(cluster))
# 
# # Placeholder for group assignation
# group_df <- 
#   expand.grid(group_number = 1:13,
#               counter = 1:8)
# group_df$remove <- FALSE
# group_df$remove[group_df$group_number > 7 &
#            group_df$counter > 5] <- TRUE
# group_df <- group_df[!group_df$remove,]
# group_df <- group_df %>%
#   dplyr::select(-remove) %>%
#   arrange(group_number)
# 
# # Loop through each group to get an expanded one
# group_df <-
#   cbind(group_df,
#         ordered_clusters)
# 
# # Get a number of clusters variable
# group_df <-
#   group_df %>%
#   group_by(group_number) %>%
#   mutate(n_clusters = max(counter)) %>%
#   # Keep only relevant variables
#   ungroup %>%
#   mutate(km = km_to_closest_health_facility) %>%
#   dplyr::select(group_number,
#                 n_clusters,
#                 health_facility,
#                 cluster,
#                 km)
```


# Recruitment

Each eligible cluster has been assigned a spray status and size, has been randomized into either "selected" (86 clusters) or not, and has been assigned to a "group" of field workers operating out of 1 or multiple health facilities. Each household within each selected cluster also is either in the "buffer" or "core", and may or may not have children.

## Table

For recruitment purposes, the following table is a list of only _eligible_ children from _eligible_ clusters, organized by group and health facility. Children living in the buffer zones have been removed. For each cluster, children are given a `random_id` which is a number between 1 and the number of (eligible) children in that cluster. This `random_id` should be used for recruitment (ie, the first child recruited should be number 1, then number 2, etc.).

The table is organized by `group_number`, so as to be of operational use. 

```{r}
final <- 
  children %>%
  # Get census data
  left_join(census_spatial_ll@data %>%
              mutate(LOCALIDADE = locality_Final,
                     BAIRRO = local_village_name) %>%
              dplyr::select(uri,
                            within_1k_voronoi_buffer_spray,
                            cluster,
                            LOCALIDADE,
                            BAIRRO,
                            lat,
                            lng),
            by = c('parent_auri' = 'uri')) %>%
  # Get group data
  left_join(clusters %>%
              filter(selected) %>%
              rename(health_facility = closest_health_facility) %>%
              dplyr::select(cluster, health_facility),
            by = 'cluster') %>%
  filter(!within_1k_voronoi_buffer_spray) %>%
  mutate(spray_status = ifelse(status, 'Spray', 'No spray')) %>%
  dplyr::select(health_facility,
                cluster,
                village_number,
                LOCALIDADE,
                BAIRRO,
                spray_status,
                house_number, keeper,
                name, dob, gender,
                permid,
                lng, lat) %>%
  filter(!is.na(health_facility))

# Randomize order and give a selection number
set.seed(639)
final <- 
  final %>%
  mutate(dummy = 1) %>%
  group_by(cluster) %>%
  mutate(random_id = cumsum(dummy)) %>%
  ungroup %>%
  dplyr::select(-dummy)

# Reorder
final <- final %>%
  arrange(health_facility, cluster, random_id)

# Make data frame
final <- data.frame(final)

# fix encoding
classes <- as.character(unlist(lapply(final, class)))

for (j in 1:ncol(final)){
  if(classes[j] == 'factor'){
    final[,j] <- as.character(final[,j])
  }
}

for (j in 1:ncol(final)){
  if(classes[j] %in% c('character', 'factor')){
    final[,j] <-
      iconv(final[,j], 'latin1', 'UTF-8')
  }
}

DT::datatable(final)
final <- data.frame(final)
write_csv(final, 'final_recruitment_table_by_group.csv')
```

## Maps

### Cluster / center maps

What follows is a map which shows all selected locations, with color indicating the health facility to which the location is assigned. The big, hollow circles are health facilities (the health facility being located at the center of the circle), and the small dots are houses selected in the random sampling pertaining to a team based at that health facility. 

Clicking on a dot reveals the house number; clicking on a circle reveals the health facility name.


```{r}
health_map <- function(){
  # Define vector of facilities
  health_facilities <- sort(unique(final$health_facility))
  
  # Define vector of colors
  colors <- colorRampPalette(brewer.pal(9, 'Spectral'))(length(unique(us$health_facility)))
  cols <- colors[as.numeric(factor(final$health_facility))]
  # Make initial map
  ll <- 
    leaflet(data = final) %>%
  addProviderTiles('Esri.WorldImagery') %>%
  addCircleMarkers(~lng,
             ~lat,
             fillOpacity= 0.2,
             opacity = 0,
             popup = ~house_number,
             color = cols,
             radius = 3)
  # Add the health facilities
  new_cols <-  colors[as.numeric(factor(us$health_facility))]
  ll <- ll %>%
    addCircleMarkers(data = us,
                     lng = us$lon,
                     lat = us$lat,
             fillOpacity= 0.3,
             opacity = 1,
             popup = us$health_facility,
             color = new_cols,
             radius = 7)
  return(ll)
  
}
health_map()
```

### Groups maps

What follows is a series of maps, one for each "group". Each map contains the locations of the children/houses assigned for recruitment to that group.

```{r}
health_facilities <- sort(us@data$health_facility)
group_map <- function(number = 1,
                      add_mop = FALSE){
  sub_data <- final %>%
    filter(health_facility == health_facilities[number])
  sub_us <- us[us@data$health_facility %in% unique(sub_data$health_facility),]
  
  color_pal <- colorRampPalette(brewer.pal(9, 'Spectral'))(length(unique(sub_data$cluster)))
  
  the_colors <- color_pal[as.numeric(factor(sub_data$cluster))]

  
  # Plot
  ll <- 
    leaflet(data = sub_us@data) %>%
  addProviderTiles('Esri.WorldImagery') %>%
  addMarkers(~lng,
             ~lat,
             popup = ~health_facility)
  if(add_mop){
    ll <- ll %>%
      addPolygons(data = mop_ll, stroke = FALSE, color = 'red')
  }
  ll <- ll  %>%
  addCircleMarkers(data = sub_data,
                     ~lng,
                     ~lat,
                     popup = as.character(sub_data$house_number),
                   radius = 3,
                   opacity = 0,
                   fillOpacity = 0.6,
                   fillColor = the_colors) %>%
    addLegend(position = 'bottomright',
              colors = 'grey',
              labels = paste0(health_facilities[number])) %>%
    addLegend(position = 'topleft',
              colors = color_pal,
              labels = levels(factor(sub_data$cluster)),
              title = 'Cluster')
  return(ll)
  
}
```


#### CENTRO DE SAUDE DE CATALE

```{r}
group_map(1)
```


#### CENTRO DE SAUDE DE CHIMUARA


```{r}
group_map(2)
```

#### CENTRO DE SAUDE DE GULAMO


```{r}
group_map(3)
```

#### CENTRO DE SAUDE DE LUALUA


```{r}
group_map(4)
```

#### CENTRO DE SAUDE DE MOPEIA


```{r}
group_map(5)
```

#### CENTRO DE SAUDE DE MUNGANE

```{r}
group_map(6)
```


#### CENTRO DE SAUDE DE NHACATUNDO


```{r}
group_map(7)
```

#### CENTRO DE SAUDE DE NOERE


```{r}
group_map(8)
```

#### CENTRO DE SAUDE  DE NZANZA


```{r}
group_map(9)
```

#### CENTRO DE SAUDE DE POSTO CAMPO


```{r}
group_map(10)
```

#### CENTRO DE SAUDE DE SANGALAZA


```{r}
group_map(11)
```

#### CENTRO DE SAUDE OITO DE MARCO

```{r}
group_map(12)
```

```{r}
# Write a master table for carlos
master_table <-
  children %>%
  # Get census data
  left_join(census_spatial_ll@data %>%
              mutate(LOCALIDADE = locality_Final,
                     BAIRRO = local_village_name) %>%
              dplyr::select(uri,
                            within_1k_voronoi_buffer_spray,
                            cluster,
                            LOCALIDADE,
                            BAIRRO,
                            lat,
                            lng),
            by = c('parent_auri' = 'uri')) %>%
  # Get group data
  left_join(final %>%
              dplyr::select(permid,
                            random_id)) %>%
  filter(!duplicated(permid))
write_csv(master_table, 'master_table_for_carlos.csv')
```

