---
title: "Cluster selection: January 2017 modifications"
author: "Joe Brew"
fig_width: 4
output:
  html_document:
    theme: flatly
    toc: yes
    toc_float: yes
  pdf_document:
    toc: yes
fig_height: 2.6
---
```{r, echo = FALSE, warning = FALSE, message = FALSE, comment = NA, error= FALSE, cache = FALSE}
# No scientific notation
options(scipen=999)

# Packages 
library(xtable)
library(knitr)
library(ggplot2) 
# library(dplyr)
library(rgeos)
library(RColorBrewer)
library(Rmisc)
# Libraries
library(rgdal)
library(raster)
library(sp)
library(leaflet)
library(ggmap)
library(readxl)
library(dplyr)
library(readr)
library(geosphere)

# Basic knitr options
opts_chunk$set(comment = NA, 
               echo = FALSE, 
               warning = FALSE, 
               message = FALSE, 
               error = TRUE, # Render report, even with errors
               cache = F)

# Helpers
source('lib/get_zambezia.R')
source('lib/helpers.R')
```

```{r setup, include=FALSE, echo = FALSE}
library(knitr)
knit_engines$set(asis = function(options) {
  if (options$echo && options$eval) knit_child(text = options$code)
})
```


Given logistical / operational challenges, we are removing clusters 119 and 58 from our selection. This means that both Mungane and Sangalaza need replacement clusters, taken from previously unselected clusters. This also means that new children will need to be randomly added for each of those clusters.

```{r, message=FALSE}
# Read in some saved objects from the December doc
# Save clusters and census for january revisions
load('selected_clusters.RData')
load('children.RData')
load('census.RData')

# Read in the master table, produced in the "health_post_assignments.Rmd"
master_table <- read_csv('master_table_for_carlos.csv')

# Read in the recruitment table, from the same document
recruitment_table <- read_csv("final_recruitment_table_by_group.csv")

# Remove anything from cluster 119
master_table <- master_table %>%
  filter(cluster != 119)
recruitment_table <-
  recruitment_table %>%
  filter(cluster != 119)

# Remove anything from cluster 58
master_table <- master_table %>%
  filter(cluster != 58)
recruitment_table <-
  recruitment_table %>%
  filter(cluster != 58)

```

## New cluster for Mungane

The below map shows the eligible clusters (originally assigned to Posto Campo, but not selected in the randomization). Click for more details on each cluster.

```{r}
# Pick a non-selected cluster from posto campo for Mungane
possible_picks <-
  clusters %>%
  filter(closest_health_facility == 'CENTRO DE SAUDE DE POSTO CAMPO',
         !clusters$selected) %>%
  mutate(x = lng,
         y = lat)
possible_picks_sp <- possible_picks
coordinates(possible_picks_sp) <- ~x+y
proj4string(possible_picks_sp) <- CRS("+proj=longlat +ellps=WGS84 +datum=WGS84")


# Generate map of possible picks
leaflet(possible_picks_sp) %>%
    addProviderTiles('Esri.WorldImagery') %>%   
  addCircleMarkers(popup = paste0('Cluster: ',
                                  possible_picks_sp$cluster))
```

The cluster numbers are `r paste0(possible_picks$cluster, collapse = ', ')`. We non-randomly choose cluster 88.

```{r}
set.seed(999)
selected <- 88 #sample(possible_picks$cluster, 1)
```

The children in the "core" area of our selected cluster (`r selected`) are below. Each child gets a number from 1 onwards, which can be used for the randomization (`random_id`).

```{r}
x <- children %>%
  dplyr::select(-house_number,
                -village_number,
                -status) %>%
  left_join(census,
            by = c('parent_auri' = 'uri')) %>%
  filter(cluster == selected) %>%
    mutate(in_core = !within_1k_voronoi_buffer_spray) %>%
  filter(in_core) %>%
  mutate(LOCALIDADE = locality_Final,
                     BAIRRO = local_village_name,
         random_id = NA) %>%
  mutate(spray_status = ifelse(status, 'Spray', 'No spray')) %>%
    mutate(health_facility = 'CENTRO DE SAUDE DE MUNGANE')
y <- x
x <- x[,c(names(recruitment_table))]
set.seed(234)
x$random_id <- sample(1:nrow(x), nrow(x), replace = FALSE)
x <- x %>% arrange(random_id)
DT::datatable(x)

# Add to the recruitment table
recruitment_table <-
  rbind(recruitment_table,
        x)
new_children <- x
```

Here is a map of the children in the new cluster.

```{r}
x_sp <- y
coordinates(x_sp) <- ~gpc_lng+gpc_lat
proj4string(x_sp) <- CRS("+proj=longlat +ellps=WGS84 +datum=WGS84")
leaflet(x_sp) %>%
    addProviderTiles('Esri.WorldImagery') %>%   
  addCircleMarkers(popup = paste0('Cluster: ',
                                  x_sp$cluster,
                                  ' | HH number: ',
                                  x_sp$household_number))
```



## New cluster for Sangalaza

The below map shows the previously non-selected for Sangalaza. Click for more details on each cluster.

```{r}

# Pick a non-selected cluster from posto campo for Mungane
possible_picks <-
  clusters %>%
  filter(closest_health_facility == 'CENTRO DE SAUDE DE SANGALAZA',
         !clusters$selected) %>%
  mutate(x = lng,
         y = lat)
possible_picks_sp <- possible_picks
coordinates(possible_picks_sp) <- ~x+y
proj4string(possible_picks_sp) <- CRS("+proj=longlat +ellps=WGS84 +datum=WGS84")


# Generate map of possible picks
# Generate map of possible picks
leaflet(possible_picks_sp) %>%
    addProviderTiles('Esri.WorldImagery') %>%   
  addCircleMarkers(popup = paste0('Cluster: ',
                                  possible_picks_sp$cluster))
```

The cluster numbers are `r paste0(possible_picks$cluster, collapse = ', ')`.

```{r}
set.seed(999)
selected <- sample(possible_picks$cluster, 1)
```

We randomly select one. It is cluster number `r selected`. The children in the "core" area of our selected cluster are below. Each child gets a number from 1 onwards, which can be used for the randomization (`random_id`).

```{r}
x <- children %>%
  dplyr::select(-house_number,
                -village_number,
                -status) %>%
  left_join(census,
            by = c('parent_auri' = 'uri')) %>%
  filter(cluster == selected) %>%
    mutate(in_core = !within_1k_voronoi_buffer_spray) %>%
  filter(in_core) %>%
  mutate(LOCALIDADE = locality_Final,
                     BAIRRO = local_village_name,
         random_id = NA) %>%
  mutate(spray_status = ifelse(status, 'Spray', 'No spray')) %>%
  mutate(health_facility = 'CENTRO DE SAUDE DE SANGALAZA')
y <- x
x <- x[,c(names(recruitment_table))]

set.seed(234)
x$random_id <- sample(1:nrow(x), nrow(x), replace = FALSE)
x <- x %>% arrange(random_id)
DT::datatable(x)

# Add to the recruitment table
recruitment_table <-
  rbind(recruitment_table,
        x)
new_children <- 
  rbind(new_children,
        x)
```

Here is a map of the children in the new cluster.

```{r}
x_sp <- y
coordinates(x_sp) <- ~gpc_lng+gpc_lat
proj4string(x_sp) <- CRS("+proj=longlat +ellps=WGS84 +datum=WGS84")
leaflet(x_sp) %>%
    addProviderTiles('Esri.WorldImagery') %>%   
  addCircleMarkers(popup = paste0('Cluster: ',
                                  x_sp$cluster,
                                  ' | HH number: ',
                                  x_sp$household_number))
```

# List of new incorporations

```{r}
new_children <- new_children %>%
  arrange(health_facility, 
          cluster,
          random_id)
DT::datatable(new_children)
write_csv(new_children, 'new_children_jan_11_modifications.csv')
```

## Master list

Having removed the problematic clusters, selected new ones, and assigned `random_id` numbers to the children in the new clusters, we can re-generate our master recruitment table (ie, the table with all core children).

```{r}
recruitment_table <- recruitment_table %>%
  arrange(health_facility, 
          cluster,
          random_id)
DT::datatable(recruitment_table)
write_csv(recruitment_table, 'recruitment_table_with_jan_11_modifications.csv')
```

