---
title: Preliminary maps and tables
author: CISM
fig_height: 2.6
fig_width: 4
output:
  html_document:
    toc: true
    toc_float: true
    theme: yeti
---
```{r, echo = FALSE, warning = FALSE, message = FALSE, comment = NA, error= FALSE, cache = FALSE}
# No scientific notation
options(scipen=999)

# Packages 
library(xtable)
library(knitr)
library(ggplot2) 
# library(dplyr)
library(rgeos)
library(RColorBrewer)
library(Rmisc)
# Libraries
library(rgdal)
library(raster)
library(sp)
library(leaflet)
library(ggmap)
library(readxl)
library(dplyr)
library(readr)

# Basic knitr options
opts_chunk$set(comment = NA, 
               echo = FALSE, 
               warning = FALSE, 
               message = FALSE, 
               error = TRUE, # Render report, even with errors
               cache = F)

```

```{r setup, include=FALSE, echo = FALSE}
library(knitr)
knit_engines$set(asis = function(options) {
  if (options$echo && options$eval) knit_child(text = options$code)
})
```


# Introduction

This document contains tables and maps pertaining to villages, clusters, and spraying status of Mopeia households. Note: this is preliminary. It uses algorithmic cleaning to remove households which were likely misclassified into the wrong village, and has not been manually checked for accuracy.

```{r}
#### READ IN AND CLEAN UP DATA

# Source script
source('lib/get_zambezia.R')

# # Read in zambezia data from odk database
# get_zambezia(get_fresh = FALSE, save = FALSE)
# # Get a more nicely named core census file
# census <- CENSUS_MOPEIA_CORE

# Read in final spreadsheet, as provided by Orvalho
# census <- read_excel('dropbox/final_joe_dataset.xlsx')

# # Read in final redcap dataset, as instructed by Charfudin 11 October 2016
# census <- readr::read_csv(file = 'redcap/CensusMopeia_DATA_2016-10-11_1541.csv')
# # Keep only relevant observations
# census <- census %>% filter(redcap_event_name == 'census_mopeia_core_arm_1')
# census <- data.frame(census)

# Read in final dataset, as given by Orvahlo on October 12, 2016
census <- read_excel('dropbox/20161012/household.xlsx')
# Get more nicely named geographic coordinate column names
census$lat <- census$y <- census$latitude <- census$gpc_lat
census$lon <- census$lng <- census$longitude <-census$x <- census$gpc_lng

# Get a shapefile for Mozambique
moz <- raster::getData('GADM', country = 'MOZ', level = 3)
moz2 <- raster::getData('GADM', country = 'MOZ', level = 2)

# Get a village number
census$village_number <- 
  as.numeric(lapply(strsplit(census$houseno_rev, split = '-'), function(x){x[2]}))

old_census <- census

# Create a spatial version of census
census_spatial <- census
coordinates(census_spatial) <- ~x+y
proj4string(census_spatial) <- proj4string(moz)

# Subset the shapefile just to Zambezia
zam <- moz[moz@data$NAME_1 == 'Zambezia',]

# Subset further to just Mopeia
mop <- moz2[moz2@data$NAME_2 == 'Mopeia',]

# Save the original proj4string
original_proj4string <- proj4string(moz)

# Prepare ggmap-compatible objects
zam_fortified <- fortify(zam, id = ID_3)
mop_fortified <- fortify(mop, id = ID_3)
zam_gg <- get_map(location = c(lon = mean(census$lon, na.rm = TRUE),
                               lat = mean(census$lat, na.rm = TRUE)),
                  maptype = 'satellite')
mop_gg <- get_map(location = c(lon = mean(census$lon, na.rm = TRUE),
                               lat = mean(census$lat, na.rm = TRUE)),
                  maptype = 'satellite')

# Switch to a projected format
# Project
census_spatial <- spTransform(census_spatial, CRS( "+init=epsg:3347" ) )
mop <- spTransform(mop, CRS("+init=epsg:3347"))
# Later, we can switch back to lat/long:
census_spatial_ll <- spTransform(census_spatial, CRS("+init=epsg:4326"))
mop_ll  <- spTransform(mop, CRS("+init=epsg:4326"))
# proj4string(mop_ll) <- proj4string(zam)

# Get the most prevalent village name for each village number
census_villages <-
  census %>%
  dplyr::group_by(village_number, village) %>% 
  tally %>%
  group_by(village_number) %>%
  summarise(correct_village_name = village[which(n == max(n))][1])
```



```{r}
# Get the area and other metrics for each village
unique_village_numbers <- sort(unique(census$village_number))
village_df <- census %>%
  group_by(village_number) %>%
  summarise(n_children = sum(number_of_children),
            houses_with_any_children = length(which(number_of_children > 0)),
            houses = n(),
            area = NA,
            incorrect_houses = NA,
            n_incorrect_houses = NA)

# Create placeholder vector of incorrect houses
incorrect_houses <- c()
  
color <- 'darkgreen'
for (i in 1:nrow(village_df)){
  message(i)
  # Get village name
  this_village <- village_df$village_number[i]
    # Get the points only
  sub_census <- 
    census_spatial[which(census_spatial$village_number == this_village),]
  # Get the border
  border <- gConvexHull(sub_census)
  # Calculate area
  if(!is.null(border)){
    this_area <- gArea(border)
    # Populate area in dataframe
    village_df$area[i] <- this_area
  }
  
  
  # Go through each household in the unit and calculate distance
  # to all other units
  distances <- gDistance(sub_census, byid = TRUE)
  median_distances <- apply(distances, 1, median)
  # Get how close they are to the closest 80% of houses in the villages
  closest_80 <- apply(distances,
                      1,
                      function(x){
                        z <- sort(x)
                        quantile(z, 0.8, na.rm = TRUE)
                      })

  # From the vector of median distances, flag those which are abnormal
  threshold <- 20000
  incorrect_house_ids <- sub_census$houseno_rev[closest_80 > threshold]
  # Stick results into dataframe
  village_df$n_incorrect_houses[i] <- length(incorrect_house_ids)
  village_df$incorrect_houses[i] <- paste0(incorrect_house_ids, collapse = ', ')
  
  # Add to the vector of incorrect houses 
  incorrect_houses <- c(incorrect_houses, incorrect_house_ids)
}
```

```{r}

# Reorder the village dataframe by size
village_df <- 
  village_df %>%
  arrange(desc(houses))

# Print table
x <- village_df %>%
  dplyr::select(-incorrect_houses)
# DT::datatable(x)

# MERGE WITH THE CLUSTER ASSIGNMENTS FROM AUGUST
cluster_assignments <- readr::read_csv('outputs/cluster_assignments.csv')
village_df <- 
  village_df %>%
  left_join(cluster_assignments)

```

<!-- All in all, `r length(incorrect_houses)` houses are suspected to be incorrectly classified (ie, either their geographic coordinates are incorrect, or their household ID number is incorrect). These are those IDs: -->

```{r}
incorrect_houses_df <- data.frame(houseno_rev = incorrect_houses)
incorrect_houses_df <-
  incorrect_houses_df %>%
  left_join(census %>%
              dplyr::select(houseno_rev, village_number)) %>%
  arrange(village_number) %>%
  dplyr::select(village_number, houseno_rev)
# DT::datatable(incorrect_houses_df)
```

<!-- All in all, `r length(which(village_df$incorrect_houses > 0))` of `r nrow(village_df)` villages appear to have incorrect household inclusions. The below maps show each of these `r length(which(village_df$incorrect_houses > 0))` villages. Note the abnormal (and large) polygon shapes, due to the incorrect inclusions. -->

```{r}
village_df$problems <- village_df$incorrect_houses > 0
village_df$row_number <- 1:nrow(village_df)
incorrect_villages <- village_df[which(village_df$incorrect_houses > 0),]
```

<!-- ## Addressing errors -->

<!-- In order to address errors, we simply remove the `r length(incorrect_houses)` houses we suspect to be incorrectly classified.   -->

```{r, eval = TRUE}
# # FILTERING OUT THOSE INCORRECT HOUSES
census <- census %>%
  filter(!houseno_rev %in% incorrect_houses)

# Create a spatial version of census
census_spatial <- census
coordinates(census_spatial) <- ~x+y
proj4string(census_spatial) <- proj4string(moz)
# Switch to a projected format
# Project
census_spatial <- spTransform(census_spatial, CRS( "+init=epsg:3347" ) )
# Later, we can switch back to lat/long:
census_spatial_ll <- spTransform(census_spatial, CRS("+init=epsg:4326"))
```


```{r, eval = TRUE}
# Remove those observations with no coordinates
census <- census %>%
  filter(!is.na(lng), !is.na(lat))

# Get centroids
village_df <-
  village_df %>%
  left_join(census %>%
    group_by(village_number) %>%
    summarise(x_centroid = mean(lng, na.rm = TRUE),
              y_centroid = mean(lat, na.rm = TRUE)))

# Remove the villages with no coordinates
village_df <- 
  village_df %>%
  filter(!is.na(x_centroid),
         !is.na(y_centroid))

# Get distances between villages
distances <- spDists(x = as.matrix(village_df %>% dplyr::select(x_centroid, y_centroid)),
        longlat = TRUE,
        diagonal = FALSE)

# Bring cluster type into census
census <- census %>% 
  left_join(village_df %>%
              dplyr::select(village_number, cluster),
            by = 'village_number')
# Create a spatial version of census
census_spatial <- census
coordinates(census_spatial) <- ~x+y
proj4string(census_spatial) <- proj4string(moz)
# Switch to a projected format
# Project
census_spatial <- spTransform(census_spatial, CRS( "+init=epsg:3347" ) )
# Later, we can switch back to lat/long:
census_spatial_ll <- spTransform(census_spatial, CRS("+init=epsg:4326"))
```


```{r}
x <- village_df %>%
  group_by(cluster) %>%
  summarise(village_numbers = paste0(village_number, collapse = ', '))

cluster_df <- census %>%
  group_by(cluster) %>%
  summarise(n_children = sum(number_of_children),
            houses_with_any_children = length(which(number_of_children > 0)),
            houses = n())

x <- x %>%
  left_join(cluster_df)
# DT::datatable(x)

```


```{r}
# Read in the spray / no-spray statuses (emailed to me)
# spray_status <- read_excel('20160817 Mopeia spray-no spray random list.xlsx')
spray_status <- read_excel('Final spray no spray list.xlsx')

# Clean up
spray_status <- spray_status[,1:11]
names(spray_status) <-
  tolower(gsub('.', '_', names(spray_status), fixed = TRUE))
names(spray_status) <- gsub(' |-', '_', names(spray_status))

# Remove NA rows
spray_status <- 
  spray_status %>%
  filter(!is.na(cluster))

# Create a status column
spray_status$status <- spray_status$spray_status
spray_status$status <-
  ifelse(is.na(spray_status$status),
         FALSE,
         ifelse(spray_status$status == 'YES',
                TRUE,
                NA))

# Create a clusters dataframe
cluster_df <- census %>%
  group_by(cluster) %>%
  summarise(n_children = sum(number_of_children),
            houses_with_any_children = length(which(number_of_children > 0)),
            houses = n())


# Join to cluster_df
cluster_df$status <- NULL
cluster_df <- 
  cluster_df %>%
  left_join(spray_status %>%
              dplyr::select(cluster,
                            status))

```

```{r}
x <- cluster_df
# DT::datatable(x %>% mutate(spray = ifelse(status, 'spray', 'no spray')) %>% dplyr::select(-status))
```

```{r}
x <- 
  census %>%
  group_by(village_number) %>%
  summarise(houses = n(),
            cluster = first(cluster)) %>%
  left_join(cluster_df %>%
              dplyr::select(cluster, status))
# DT::datatable(x %>% mutate(spray = ifelse(status, 'spray', 'no spray')) %>% dplyr::select(-status))
# Write a csv
# readr::write_csv(x, '~/Desktop/spray_no_spray_assignments.csv')
# Get spray status per row
census <- left_join(
  census,
  cluster_df %>%
    dplyr::select(cluster, status)
)
# Create a spatial version of census
census_spatial <- census
coordinates(census_spatial) <- ~x+y
proj4string(census_spatial) <- proj4string(moz)
# Switch to a projected format
# Project
census_spatial <- spTransform(census_spatial, CRS( "+init=epsg:3347" ) )
# Later, we can switch back to lat/long:
census_spatial_ll <- spTransform(census_spatial, CRS("+init=epsg:4326"))


village_list <- list()
for (i in 1:nrow(village_df)){
  village_list[[i]] <- village_df[i,]
}
```


# Table

The below table shows each household in the dataset. It is interactive and searchable. In the cases in which a `village_number` has more than one `local_village_name`, this indicates a discrepancy in the raw data.

```{r}
x <- census %>%
  dplyr::select(status,
                cluster,
                village_number,
                houseno_rev,
                local_village_name,
                locality,
                administrative_post,
                health_facility) %>%
  mutate(status = ifelse(is.na(status), 'Undetermined', ifelse(status, 'Spray', 'No spray'))) %>%
  arrange(village_number, houseno_rev)
x$local_village_name <- stringr::str_replace_all(x$local_village_name,"[^[:graph:]]", " ")

# DT::datatable(x)
# 
# # Write table for joe wagman
# x <- left_join(old_census %>% 
#                  dplyr::select(houseno_rev),
#                x %>%
#                  mutate(problematic = FALSE)) %>%
#   mutate(problematic = ifelse(is.na(problematic), TRUE, problematic)) 
# x <- x %>% arrange(village_number, houseno_rev)
# write_csv(x, '~/Desktop/table_for_wagman.csv')

# Get a table for the October 17 request
x <- census %>%
  dplyr::select(status,
                cluster,
                village_number,
                houseno_rev,
                village,
                local_village_name,
                other_village,
                locality,
                administrative_post,
                health_facility,
                head_name,
                lat,
                lng) %>%
  mutate(status = ifelse(is.na(status), 'Undetermined', ifelse(status, 'Spray', 'No spray'))) %>%
  
  arrange(village_number, houseno_rev)
x$local_village_name <- stringr::str_replace_all(x$local_village_name,"[^[:graph:]]", " ")


# Write table for edgar
x <- left_join(old_census %>% 
                 dplyr::select(houseno_rev),
               x %>%
                 mutate(problematic = FALSE)) %>%
  mutate(problematic = ifelse(is.na(problematic), TRUE, problematic)) 
x <- x %>% arrange(village_number, houseno_rev)

# Keep only relevant variables
x <- x %>%
  dplyr::select(administrative_post,
                locality,
                local_village_name,
                houseno_rev,
                head_name,
                status,
                problematic,
                lat,
                lng) %>%
  rename(`Administrative name` = administrative_post,
         `Locality name` = locality,
         `Local village name` = local_village_name,
         `Revised house number` = houseno_rev,
         `Spray status` = status,
         `Head of household` = head_name,
         Latitude = lat,
         Longitude = lng)
write_csv(x, '~/Desktop/rose_lourdes_table.csv')

# Also make a csv of each village_number with village names
# x <- old_census %>%
#   group_by(village_number, local_village_name) %>%
#   tally %>%
#   arrange(village_number, desc(n)) %>%
#   rename(frequency = n)
# write_csv(x, '~/Desktop/table_of_village_names.csv')
```

# Maps

```{r}

# CREATE FUNCTION FOR LEAFLET MAP OF VILLAGE
leaflet_village <- function(i){
  
    this_village_number <- village_df$village_number[i]
   # Get the points only
  sub_census <- census_spatial_ll[which(census_spatial_ll$village_number == this_village_number),]
  # Get the border
  border <- gConvexHull(sub_census)
  
  the_data <- sub_census
  the_border <- border
  
  # Generate colors
  colors <- rep('blue', nrow(the_data))
  colors[the_data$status] <- 'red'
  
  # The village number
  the_village_number <- the_data$village_number[1]
  
  # The village df row
  the_village_df_row <- village_df %>% filter(village_number == the_village_number)
  
  #   # Legend title
  legend_title <- paste0('Village: ',
           the_village_number)
  

  ll <- 
  leaflet() %>%
  addProviderTiles("OpenStreetMap.Mapnik") %>%
  addCircleMarkers(lng = the_data$lng,
    lat = the_data$lat,
    color = colors,
    fillColor = colors,
    radius = 2.5,
    opacity = 0,
    fillOpacity = 0.5,
    popup = paste0('Household: ', the_data$houseno_rev, ' Village number: ',
    the_data$village_number)) %>%
    addLegend("bottomright", 
              colors = c('red', 'blue'),
              labels = c(paste0('Spray'), 
                         'No spray'),
    title = legend_title,
    opacity = 1 )
  
  if(class(border)[[1]] != 'SpatialPoints'){
    ll <-
      ll %>%
          addPolylines(data = the_border, color = 'black',
                 dashArray = '1,5')
  }
  return(ll)
}



```

## Master static map (all villages in one map)

The below map shows each household (point) with each village's boundary (polygon). Colors are random, but each village gets only one color. This map was produced after removing suspicious/incorrectly coded houses (ie, those that were far from their supposed village.)

```{r}
plot(mop_ll)
cols <- rainbow(nrow(village_df))
for (i in 1:nrow(village_df)){
  
    this_village_number <- village_df$village_number[i]
   # Get the points only
  sub_census <- census_spatial_ll[which(census_spatial_ll$village_number == this_village_number),]
  # Get the border
  border <- gConvexHull(sub_census)
  
  the_data <- sub_census
  the_border <- border
  
  plot(the_border, add = TRUE,
       col = adjustcolor(cols[i], alpha.f = 0.2),
       border =  adjustcolor(cols[i], alpha.f = 0.3))
  points(the_data, pch = '.',
         col = adjustcolor(cols[i], alpha.f = 0.6))
}
```

## Interactive maps (one per village)

The below maps show each village. Households suspected to be misclassified into the wrong village have been removed. Because this removal was done algorithmically (and relatively quickly), it has not been fine-tuned (hence, the presence of some houses which are very far from the others in the village). These are preliminary maps, and should be understood as such.


```{r}
village_df$row_number <- 1:nrow(village_df)
maps <- lapply(village_df$row_number, function(x){
  leaflet_village(x)
})

htmltools::tagList(maps)
# for(i in 1:3){
# # for(i in 1:nrow(village_df)){
#   this_village_number <- village_df$village_number[i]
#    # Get the points only
#   sub_census <- census_spatial_ll[which(census_spatial_ll$village_number == this_village_number),]
#   # Get the border
#   border <- gConvexHull(sub_census)
# 
#   leaflet_village(the_data = sub_census,
#                   the_border = border) 
#   
#   #   # Plot Mopeia
#   # plot(mop_ll)
#   # # Plot the border
#   # plot(border, add = TRUE, col = adjustcolor(color, alpha.f = 0.3),
#   #      border = adjustcolor(color, alpha.f = 0.8))
#   # # Add the points
#   # points(sub_census,
#   #        pch = '.',
#   #        col = adjustcolor(color, alpha.f = 0.6))
#   # # Add title
#   # title(main = paste0('Village number: ', unique_village_numbers[i]))
# }
```


