---
title: "Maps with buffers"
author: "Eldo Elobolobo, Orvahlo Augusto, Joe Brew, Carlos Chaccour, Edgar Jamisse, Charfudin Sacoor, Francisco Saute"
fig_width: 4
output:
  html_document:
    theme: flatly
    toc: yes
    toc_float: yes
  pdf_document:
    toc: yes
fig_height: 2.6
---
```{r, echo = FALSE, warning = FALSE, message = FALSE, comment = NA, error= FALSE, cache = FALSE}
# No scientific notation
options(scipen=999)

# Packages 
library(xtable)
library(knitr)
library(ggplot2) 
# library(dplyr)
library(rgeos)
library(RColorBrewer)
library(Rmisc)
# Libraries
library(rgdal)
library(raster)
library(sp)
library(leaflet)
library(ggmap)
library(readxl)
library(dplyr)
library(readr)
library(geosphere)

# Basic knitr options
opts_chunk$set(comment = NA, 
               echo = FALSE, 
               warning = FALSE, 
               message = FALSE, 
               error = TRUE, # Render report, even with errors
               cache = F)

# Helpers
source('lib/get_zambezia.R')
source('lib/helpers.R')
```

```{r setup, include=FALSE, echo = FALSE}
library(knitr)
knit_engines$set(asis = function(options) {
  if (options$echo && options$eval) knit_child(text = options$code)
})
```


```{r}
#### READ IN AND CLEAN UP DATA

# Read in final dataset, as given by Orvahlo on October 24, 2016
# census <- read_excel('Limpesa_Final__241016_For maps.xlsx')
# Read in corrected final dataset, sent by Edgar on October 30, 2016
# census <- read_excel('Casos problematticos resolvidos_Edgar Jamisse_Mopeia_2016_28_10.xls')
census <- read_csv('Casos problematticos resolvidos_Edgar Jamisse_Mopeia_2016_28_10.csv')
# census <- read_csv('CensusMopeia_DATA_2016-11-11_1355.csv')
# Fix weird header
names(census)[6] <- 'locality_Final'

# Get more nicely named geographic coordinate column names
census$lat <- census$y <- census$latitude <- census$gpc_lat
census$lon <- census$lng <- census$longitude <-census$x <- census$gpc_lng

# Get a shapefile for Mozambique
moz <- raster::getData('GADM', country = 'MOZ', level = 3)
moz2 <- raster::getData('GADM', country = 'MOZ', level = 2)

# GET CORRECT FINAL COLUMNS, AS EXPLAINED IN EMAIL
# Get a village number
census$village_number <- census$`village number_final`
census$administrative_post <- census$administrative_post_final
census$house_number <- census$houseno_Final_1
census$locality <- census$locality_Final

# Convert village number to numeric
census$village_number <- 
  as.numeric(as.character(census$village_number))

# Get the corrected locality and bairro names sent by edgar on nov 7
locality_corrections <- read_excel('Mapa Actualizado de bairros Censados_07_11-2016_Edgar-Revisao_FIN.xlsx', skip = 2)

# Join to the census
x <- 
  census %>%
  left_join(locality_corrections,
            by = c('village_number' = 'CODIGO')) %>%
  mutate(BAIRRO = toupper(BAIRRO))
# Make updates
x <- x %>%
  mutate(administrative_post = 
           ifelse(administrative_post != `POSTO ADMIN` & !is.na(`POSTO ADMIN`), 
                  `POSTO ADMIN`, administrative_post),
         locality = ifelse(locality != LOCALIDADE & !is.na(LOCALIDADE), 
                           LOCALIDADE, locality),
         local_village_name = ifelse(local_village_name != BAIRRO & !is.na(BAIRRO),
                         BAIRRO, local_village_name)) %>%
  dplyr::select(-`POSTO ADMIN`, LOCALIDADE, BAIRRO)
# Overwite census
census <- x
rm(x)
# Remove all NA rows
census <- census %>%
  filter(!is.na(lng),
         !is.na(lat),
         !is.na(village_number),
         !is.na(house_number))


# Save old version of census
census_old <- census

# Create a spatial version of census
census_spatial <- census
coordinates(census_spatial) <- ~x+y
proj4string(census_spatial) <- proj4string(moz)

# Subset the shapefile just to Zambezia
zam <- moz[moz@data$NAME_1 == 'Zambezia',]

# Subset further to just Mopeia
mop <- moz2[moz2@data$NAME_2 == 'Mopeia',]

# Save the original proj4string
original_proj4string <- proj4string(moz)

# Prepare ggmap-compatible objects
zam_fortified <- fortify(zam, id = ID_3)
mop_fortified <- fortify(mop, id = ID_3)
zam_gg <- get_map(location = c(lon = mean(census$lon, na.rm = TRUE),
                               lat = mean(census$lat, na.rm = TRUE)),
                  maptype = 'satellite')
mop_gg <- get_map(location = c(lon = mean(census$lon, na.rm = TRUE),
                               lat = mean(census$lat, na.rm = TRUE)),
                  maptype = 'satellite')

# Switch to a projected format
# Project
census_spatial <- spTransform(census_spatial, CRS( "+init=epsg:3347" ) )
mop <- spTransform(mop, CRS("+init=epsg:3347"))
# Later, we can switch back to lat/long:
census_spatial_ll <- spTransform(census_spatial, CRS("+init=epsg:4326"))
mop_ll  <- spTransform(mop, CRS("+init=epsg:4326"))
# proj4string(mop_ll) <- proj4string(zam)
```

```{r}
# Read in data at the child-specific level
children <- read_csv('redcap/CensusMopeia_DATA_2016-11-01_1734.csv')

# Keep only those households with children
children <- 
  children %>%
  filter(!is.na(perm_id_final)) %>%
  mutate(permid = perm_id_final) %>%
  dplyr::select(-perm_id_final)

# Keep only those children which appear in the non-problematic census records
children <-
  children %>%
  filter(parent_auri %in% census$uri) 

# Keep only the columns of interest
children <- 
  children %>%
  dplyr::select(permid,
                dob,
                name,
                gender,
                keeper,
                parent_auri)

# Get the household id
children <-
  children %>%
  left_join(census %>% 
              dplyr::select(house_number,
                            village_number,
                            uri),
            by = c('parent_auri' = 'uri'))
```


```{r}
if('clean_maps_with_buffers_data.RData' %in% dir()){
  load('clean_maps_with_buffers_data.RData')
} else {
  # Get spray status (will use for clusters too)
  get_spray_status() # produces spray_village (village level) and spray_status (cluster level)
  
  # Make a dataframe of just villages (with cluster assignments)
  village_df <- make_village_df(census = census,
                                spray_village = spray_village)
  
  # Get a distance matrix (takes 2-3 minutes)
  distance_matrix <- get_distance_matrix(spatial_ll_census = census_spatial_ll)
  
  # Identify problems
  identify_problems()
  census$problematic <- census$house_number %in% incorrect_houses
  census_spatial$problematic <- census_spatial$house_number %in% incorrect_houses
  census_spatial_ll$problematic <- census_spatial_ll$house_number %in% incorrect_houses
  
  # SUBSET TO REMOVE PROBLEMATIC
  # Get distance matrix after subset
  distance_matrix_old <- distance_matrix
  distance_matrix <- distance_matrix[!census$problematic,
                                         !census$problematic]
  
  # Identify the nearest neighbors of each house
  nn <- nearest_neighbor(census_spatial_ll = census_spatial_ll,
                             distance_matrix_old = distance_matrix_old,
                             n = 5)
  census$nearest_neighbors <- nn
  census_old$nearest_neighbors <- nn
  census_spatial$nearest_neighbors <- nn
  census_spatial_ll$nearest_neighbors <- nn
  
  # Remove incorrect houses
  census <- census[which(!census$house_number %in% incorrect_houses),]
  census_spatial <- census_spatial[which(!census_spatial$house_number %in% incorrect_houses),]
  census_spatial_ll <- census_spatial_ll[which(!census_spatial_ll$house_number %in% incorrect_houses),]
  
  village_df <- village_df %>% filter(!is.na(village_number),
                                      !duplicated(village_number))

  # Join spray status to objects
  joiner <- 
    village_df %>%
                dplyr::select(village_number, cluster, status)
  
  census_spatial_ll@data <-
    left_join(census_spatial_ll@data,
              joiner,
              by = 'village_number')
  census_spatial@data <-
    left_join(census_spatial@data,
              joiner,
              by = 'village_number')
  census <-
    left_join(census,
              joiner,
              by = 'village_number')
  
  # Get the voronoi tesselation surface
  vv <- voronoi(spatial_census_ll = census_spatial_ll)
  # Get the voronoi tesselation surface at the level of the spray zones
  vv_spray <- voronoi_spray_level(spatial_census_ll = census_spatial_ll)
  
  # Get a non lat long version of voronoi surface
  # For the village level
  proj4string(vv) <- proj4string(census_spatial_ll)
  vv_ll <- vv
  vv <- sp::spTransform(vv, CRSobj = proj4string(census_spatial))
  
  # For the spray-level
  proj4string(vv_spray) <- proj4string(census_spatial_ll)
  vv_spray_ll <- vv_spray
  vv_spray <- sp::spTransform(vv_spray, CRSobj = proj4string(census_spatial))
    
  # Get the buffered voronoi tesselation surface
  vvb <- rgeos::gBuffer(vv,  # village level
                        byid = TRUE, 
                        id = vv$village_number,
                        width = -1000)
  vvb_spray <- rgeos::gBuffer(vv_spray,  # village level
                        byid = TRUE, 
                        id = vv$status,
                        width = -1000)
  
  # Convert back to lat lon
  vvb_ll <- spTransform(vvb, CRSobj = proj4string(census_spatial_ll))
  vvb_ll_spray <- spTransform(vvb_spray, CRSobj = proj4string(census_spatial_ll))
  
  # Identify buffers voronoi
  census_spatial_ll <- 
    identify_buffers_voronoi(
    census_spatial_ll = census_spatial_ll,
           distance_matrix = distance_matrix,
           vvb_ll = vvb_ll)
  
  # Identify buffers voronoi at spray level
  census_spatial_ll <- 
    identify_buffers_voronoi_spray_level(census_spatial_ll = census_spatial_ll,
           distance_matrix = distance_matrix,
           vvb_ll_spray = vvb_ll_spray)
  
  # Get buffers through distance to other villages
  census_spatial_ll <- identify_buffers(spatial_ll_census = census_spatial_ll,
                       distances = distance_matrix)
  
  # Update other objects with variables only in census_spatial_ll
  update_other_objects()

  # Save
  save.image('clean_maps_with_buffers_data.RData')  
}


# Set all NA statuses to FALSE
census$status[is.na(census$status)] <- FALSE
census_spatial$status[is.na(census_spatial$status)] <- FALSE
census_spatial_ll$status[is.na(census_spatial_ll$status)] <- FALSE
village_df$status[is.na(village_df$status)] <- FALSE

# Resource helpers
source('lib/helpers.R')

# Get children's spray status
children <- 
  left_join(children,
            census %>% 
              dplyr::select(uri,
                            status),
            by = c('parent_auri' = 'uri'))
# Remove those with no spray status
children <- 
  children %>%
  filter(!is.na(status))
```


# Introduction

This document was produced originally on Saturday, September 9, 2017. Its structure is based on the `clean_maps_with_buffers.html` (from November 2016), but with updates to take into account discrepancies between assigned versus real spray statuses. It relies on the database, excel assignment sheets, a `.csv` of in-the-field spray status (ie, "real" spray statuses) sent by Eldo in early September 2017 ("field_status.csv"). Additionally, it uses data sent from Charfudin Sacoor to Joe Brew on the evening of Sunday, October 30th, 2016 (`Casos problematticos resolvidos_Edgar Jamisse_Mopeia_2016_28_10.xlsx`).


# Concordance between assigned and real spray statuses

## Table  

The below table show sthe spray assignments and real sprayings.  Variable meanings are as follows:  

- `original_assignment` means the spray status per the original assignment.  
- `assigned` means the spray status per Eldo's table (ie, `ABT(Spray_list)` in the table sent by Eldo). In theory this should match perfectly with `original_assignment`: it doesn't.  
- `real` is the actual spray status, ie the `CISM (Random_List)` column in the table sent by Eldo.  

```{r}
# Read in data sent from eldo
eldo <- read_csv('field_status.csv') %>%
  mutate(village_number = as.numeric(Village_id)) %>%
  rename(assigned = `ABT(Spray_list)`,
         real = `CISM (Random_List)`) %>%
  dplyr::select(village_number, assigned, real)

# Join to village df
village_df <-
  left_join(village_df,
            eldo)

village_df <-
  village_df %>%
  mutate(original_assignment = ifelse(status, 'Spray', 'No_Spray'))

DT::datatable(village_df %>%
                dplyr::select(village_number,
                              original_assignment,
                              assigned,
                              real))
```

Because of the discrepancy between `assigned` (Eldo) and `original_assignment` (Joe), I'm going to assume that `assigned` is correct, and that `original_assignment` is incorrect.

## Contingency table  

```{r}
concordance <- left_join(x = census,
                         y = village_df,
                         by = 'village_number')

table(paste0('Assigned ', concordance$assigned), 
      paste0('Received ', concordance$real))
```

## Maps

The below map has one point for each household. The color indicates whether its status is "OK" (got what it should have) or "Mismatch".

```{r}


# Get a color code
concordance <- 
  concordance %>%
  mutate(outcome = ifelse(assigned == 'Spray' &
                          real == 'Spray', 'OK: Spray',
                        ifelse(assigned == 'Spray' &
                                 real != 'Spray', 'Mismatch: should have gotten sprayed but did not',
                               ifelse(assigned != 'Spray' &
                                        real == 'Spray', 'Mismatch: should not have gotten sprayed but did',
                                      ifelse(assigned != 'Spray' &
                                               real != 'Spray', 'OK: No spray', 
                                             'No assignment or outcome')))))


ggplot() +
  geom_polygon(data = mop_fortified,
               aes(x = long,
                   y = lat,
                   group = group)) +
  geom_point(data = concordance,
             aes(x = longitude,
                 y = latitude,
                 color = outcome),
             alpha = 0.7,
             size = 0.3) +
  ggthemes::theme_map() +
  theme(legend.position = 'bottom',
        legend.text = element_text(size = 8)) +
  
  guides(color=guide_legend(ncol=3))
```

## Re-tesselation  

Our original buffers are no longer valid, given the presence of spraying in no-spray zones, and the presence of non-sprayed houses in spray zones. We therefore re-draw boundaries and buffers based on the "real" status of each household.

The below map shows the "real" spray vs no spray zones

```{r}
concordance_spatial <- concordance
# Overwrite status with new status
concordance_spatial$status <- concordance_spatial$real
concordance_spatial$status <- ifelse(concordance_spatial$status == 'Spray', TRUE, FALSE)
coordinates(concordance_spatial) <- ~longitude+latitude
xx <- voronoi_spray_level(spatial_census_ll = concordance_spatial)
status <- xx$status

plot(xx, col = adjustcolor(ifelse(xx$status, 'darkred', 'darkgreen'), alpha.f = 0.6), border = adjustcolor('black', alpha.f = 0.5))
legend('bottomright',
       fill = adjustcolor(c('darkred', 'darkgreen'), alpha.f = 0.6),
       legend = c('Spray', 'No spray'))
```

Compare that to our original assignment below.

```{r}
the_data <- vv_spray_ll
the_colors <- ifelse(the_data@data$status, 'darkred', 'darkgreen')
the_colors <- adjustcolor(the_colors, alpha.f = 0.6)
status <- the_data$status
plot(mop_ll)
plot(the_data, add = TRUE, col = the_colors, border = adjustcolor('black', alpha.f = 0.5))
legend('bottomright',
       fill = the_colors,
       legend = c('Spray', 'No spray'))
```

We need to buffer the "real" assignments. The result is as follows:

```{r}
proj4string(concordance_spatial) <-
  proj4string(moz)
# For the spray-level
xx_spray <- xx
proj4string(xx_spray) <- proj4string(moz)
xx_spray_ll <- xx_spray
xx_spray <- sp::spTransform(xx_spray, CRSobj = proj4string(vvb))
  
# Get the buffered voronoi tesselation surface
xxb_spray <- rgeos::gBuffer(xx_spray, 
                      byid = TRUE, 
                      id = xx$status,
                      width = -1000)

# Convert back to lat lon
xxb_ll_spray <- spTransform(xxb_spray, CRSobj = proj4string(concordance_spatial))

plot(mop_ll)
plot(xxb_ll_spray, col = adjustcolor(ifelse(status, 'darkred', 'darkgreen'), alpha.f = 0.6), border = adjustcolor('black', alpha.f = 0.5),
     add = TRUE)
legend('bottomright',
       fill = adjustcolor(c('darkred', 'darkgreen'), alpha.f = 0.6),
       legend = c('Spray', 'No spray'))
```

Compare that to our original buffered spray assignments below.


```{r}
vvb_ll_spray <- spTransform(vvb_ll_spray,
                            CRSobj = proj4string(xxb_ll_spray))
the_data <- vvb_ll_spray
the_colors <- ifelse(the_data@data$status, 'darkred', 'darkgreen')
the_colors <- adjustcolor(the_colors, alpha.f = 0.6)
status <- the_data$status
plot(mop_ll)
plot(the_data, col = the_colors, border = adjustcolor('black', alpha.f = 0.5), add = TRUE)
legend('bottomright',
       fill = the_colors,
       legend = c('Spray', 'No spray'))
```

## Spatial overlay

The below map shows the overlay of both the original assignments and the real results. The original assignments are shown in diagonal lines, whereas the reality is shown in fill.

```{r}
plot(vvb_ll_spray,
     col = adjustcolor(ifelse(status, 'darkred', 'darkgreen'), alpha.f = 0.9),
     angle = 45,
     density = 20)
plot(xxb_ll_spray,
     add = TRUE,
     col = adjustcolor(ifelse(status, 'darkred', 'darkgreen'), alpha.f = 0.5))
```

## Sub-overlay

### Spray zones

```{r}

xx_spray <- xxb_ll_spray[xxb_ll_spray$status,]
vv_spray <- vvb_ll_spray[vvb_ll_spray$status,]
xx_no_spray <- xxb_ll_spray[!xxb_ll_spray$status,]
vv_no_spray <- vvb_ll_spray[!vvb_ll_spray$status,]

par(mfrow = c(1,2))

plot(vv_spray, col = 'red', main = 'Assigned')
plot(xx_spray, col = 'red', main = 'Real')
par(mfrow = c(1,1))
```

### Non-spray zones

```{r}
par(mfrow = c(1,2))

plot(vv_no_spray, col = 'green', main = 'Assigned')
plot(xx_no_spray, col = 'green', main = 'Real')
par(mfrow = c(1,1))
```

### Assigned spray, received no spray

```{r}
mismatch <- gDifference(xx_no_spray,
                           vv_no_spray)
plot(mop_ll)
plot(mismatch, col = 'blue', add = TRUE)
```

### Assigned no spray, received spray

```{r}
mismatch <- gDifference(xx_spray,
                           vv_spray)
plot(mop_ll)
plot(mismatch, col = 'blue', add = TRUE)
```


# Further data products

Given the above overlays, we can define which households are in original vs. real spray/no-spray/buffer zones. These data will be saved and sent separately in machine-readable format (ie, `.csv`). 

```{r}
concordance %>%
  dplyr::select(house_number,
                assigned,
                real) %>%
  mutate(concordance = ifelse(assigned == real, 'OK', 'NOT OK')) %>%
  sample_n(100) %>%
  
  DT::datatable()
```

# Code 

Full code at https://github.com/joebrew/zambezia.


