---
title: Operational helper document
author: CISM
fig_height: 2.6
fig_width: 4
output:
  html_document:
    toc: true
    toc_float: true
    theme: yeti
---
```{r, echo = FALSE, warning = FALSE, message = FALSE, comment = NA, error= FALSE, cache = FALSE}
# No scientific notation
options(scipen=999)

# Packages 
library(xtable)
library(knitr)
library(ggplot2) 
# library(dplyr)
library(rgeos)
library(RColorBrewer)
library(Rmisc)
# Libraries
library(rgdal)
library(raster)
library(sp)
library(leaflet)
library(ggmap)
library(readxl)
library(dplyr)
library(readr)

# Basic knitr options
opts_chunk$set(comment = NA, 
               echo = FALSE, 
               warning = FALSE, 
               message = FALSE, 
               error = TRUE, # Render report, even with errors
               cache = F)

```

```{r setup, include=FALSE, echo = FALSE}
library(knitr)
knit_engines$set(asis = function(options) {
  if (options$echo && options$eval) knit_child(text = options$code)
})
```


```{r}
#### READ IN AND CLEAN UP DATA

# Source script
source('lib/get_zambezia.R')

# # Read in zambezia data from odk database
# get_zambezia(get_fresh = FALSE, save = FALSE)
# # Get a more nicely named core census file
# census <- CENSUS_MOPEIA_CORE

# Read in final spreadsheet, as provided by Orvalho
# census <- read_excel('dropbox/final_joe_dataset.xlsx')

# # Read in final redcap dataset, as instructed by Charfudin 11 October 2016
# census <- readr::read_csv(file = 'redcap/CensusMopeia_DATA_2016-10-11_1541.csv')
# # Keep only relevant observations
# census <- census %>% filter(redcap_event_name == 'census_mopeia_core_arm_1')
# census <- data.frame(census)

# Read in final dataset, as given by Orvahlo on October 12, 2016
census <- read_excel('dropbox/20161012/household.xlsx')

# Get more nicely named geographic coordinate column names
census$lat <- census$y <- census$latitude <- census$gpc_lat
census$lon <- census$lng <- census$longitude <-census$x <- census$gpc_lng

# Get a shapefile for Mozambique
moz <- raster::getData('GADM', country = 'MOZ', level = 3)
moz2 <- raster::getData('GADM', country = 'MOZ', level = 2)

# Get a village number
census$village_number <- 
  as.numeric(lapply(strsplit(census$houseno_rev, split = '-'), function(x){x[2]}))

# Create a spatial version of census
census_spatial <- census
coordinates(census_spatial) <- ~x+y
proj4string(census_spatial) <- proj4string(moz)

# Subset the shapefile just to Zambezia
zam <- moz[moz@data$NAME_1 == 'Zambezia',]

# Subset further to just Mopeia
mop <- moz2[moz2@data$NAME_2 == 'Mopeia',]

# Save the original proj4string
original_proj4string <- proj4string(moz)

# Prepare ggmap-compatible objects
zam_fortified <- fortify(zam, id = ID_3)
mop_fortified <- fortify(mop, id = ID_3)
zam_gg <- get_map(location = c(lon = mean(census$lon, na.rm = TRUE),
                               lat = mean(census$lat, na.rm = TRUE)),
                  maptype = 'satellite')
mop_gg <- get_map(location = c(lon = mean(census$lon, na.rm = TRUE),
                               lat = mean(census$lat, na.rm = TRUE)),
                  maptype = 'satellite')

# Switch to a projected format
# Project
census_spatial <- spTransform(census_spatial, CRS( "+init=epsg:3347" ) )
mop <- spTransform(mop, CRS("+init=epsg:3347"))
# Later, we can switch back to lat/long:
census_spatial_ll <- spTransform(census_spatial, CRS("+init=epsg:4326"))
mop_ll  <- spTransform(mop, CRS("+init=epsg:4326"))
# proj4string(mop_ll) <- proj4string(zam)

# Get the most prevalent village name for each village number
census_villages <-
  census %>%
  dplyr::group_by(village_number, village) %>% 
  tally %>%
  group_by(village_number) %>%
  summarise(correct_village_name = village[which(n == max(n))][1])
```



```{r}
# Get the area and other metrics for each village
unique_village_numbers <- sort(unique(census$village_number))
village_df <- census %>%
  group_by(village_number) %>%
  summarise(n_children = sum(number_of_children),
            houses_with_any_children = length(which(number_of_children > 0)),
            houses = n(),
            area = NA,
            incorrect_houses = NA,
            n_incorrect_houses = NA)

# Create placeholder vector of incorrect houses
incorrect_houses <- c()
  
color <- 'darkgreen'
for (i in 1:nrow(village_df)){
  message(i)
  # Get village name
  this_village <- village_df$village_number[i]
    # Get the points only
  sub_census <- 
    census_spatial[which(census_spatial$village_number == this_village),]
  # Get the border
  border <- gConvexHull(sub_census)
  # Calculate area
  if(!is.null(border)){
    this_area <- gArea(border)
    # Populate area in dataframe
    village_df$area[i] <- this_area
  }
  
  
  # Go through each household in the unit and calculate distance
  # to all other units
  distances <- gDistance(sub_census, byid = TRUE)
  median_distances <- apply(distances, 1, median)

  # From the vector of median distances, flag those which are abnormal
  threshold <- 20000
  incorrect_house_ids <- sub_census$houseno_rev[median_distances > threshold]
  # Stick results into dataframe
  village_df$n_incorrect_houses[i] <- length(incorrect_house_ids)
  village_df$incorrect_houses[i] <- paste0(incorrect_house_ids, collapse = ', ')
  
  # Add to the vector of incorrect houses 
  incorrect_houses <- c(incorrect_houses, incorrect_house_ids)
}

# Reorder the village dataframe by size
village_df <- 
  village_df %>%
  arrange(desc(houses))

# Print table
x <- village_df %>%
  dplyr::select(-incorrect_houses)
# DT::datatable(x)

# MERGE WITH THE CLUSTER ASSIGNMENTS FROM AUGUST
cluster_assignments <- readr::read_csv('outputs/cluster_assignments.csv')
village_df <- 
  village_df %>%
  left_join(cluster_assignments)

```

<!-- All in all, `r length(incorrect_houses)` houses are suspected to be incorrectly classified (ie, either their geographic coordinates are incorrect, or their household ID number is incorrect). These are those IDs: -->

```{r}
incorrect_houses_df <- data.frame(houseno_rev = incorrect_houses)
incorrect_houses_df <-
  incorrect_houses_df %>%
  left_join(census %>%
              dplyr::select(houseno_rev, village_number)) %>%
  arrange(village_number) %>%
  dplyr::select(village_number, houseno_rev)
# DT::datatable(incorrect_houses_df)
```

<!-- All in all, `r length(which(village_df$incorrect_houses > 0))` of `r nrow(village_df)` villages appear to have incorrect household inclusions. The below maps show each of these `r length(which(village_df$incorrect_houses > 0))` villages. Note the abnormal (and large) polygon shapes, due to the incorrect inclusions. -->

# Villages

The below maps show each village, as well as the "convex hull" (ie, boundaries) of that village's area.

```{r, eval = TRUE}
incorrect_villages <- village_df[which(village_df$incorrect_houses > 0),]
for(i in 1:nrow(incorrect_villages)){
  this_village_number <- incorrect_villages$village_number[i]
   # Get the points only
  sub_census <- census_spatial_ll[which(census_spatial_ll$village_number == this_village_number),]
  # Plot Mopeia
  plot(mop_ll)
  # Get the border
  border <- gConvexHull(sub_census)
  # Plot the border
  plot(border, add = TRUE, col = adjustcolor(color, alpha.f = 0.3),
       border = adjustcolor(color, alpha.f = 0.8))
  # Add the points
  points(sub_census,
         pch = '.',
         col = adjustcolor(color, alpha.f = 0.6))
  # Add title
  title(main = paste0('Village number: ', unique_village_numbers[i]))
}
```

<!-- ## Addressing errors -->

<!-- In order to address errors, we simply remove the `r length(incorrect_houses)` houses we suspect to be incorrectly classified.   -->

```{r}
# # NOT FILTERING OUT THOSE INCORRECT HOUSES
census <- census %>%
  filter(!houseno_rev %in% incorrect_houses)

# Create a spatial version of census
census_spatial <- census
coordinates(census_spatial) <- ~x+y
proj4string(census_spatial) <- proj4string(moz)
# Switch to a projected format
# Project
census_spatial <- spTransform(census_spatial, CRS( "+init=epsg:3347" ) )
# Later, we can switch back to lat/long:
census_spatial_ll <- spTransform(census_spatial, CRS("+init=epsg:4326"))
```


```{r}
# Remove those observations with no coordinates
census <- census %>%
  filter(!is.na(lng), !is.na(lat))

# Get centroids
village_df <-
  village_df %>%
  left_join(census %>%
    group_by(village_number) %>%
    summarise(x_centroid = mean(lng, na.rm = TRUE),
              y_centroid = mean(lat, na.rm = TRUE)))

# Remove the villages with no coordinates
village_df <- 
  village_df %>%
  filter(!is.na(x_centroid),
         !is.na(y_centroid))

# Get distances between villages
distances <- spDists(x = as.matrix(village_df %>% dplyr::select(x_centroid, y_centroid)),
        longlat = TRUE,
        diagonal = FALSE)

# Bring cluster type into census
census <- census %>% 
  left_join(village_df %>%
              dplyr::select(village_number, cluster),
            by = 'village_number')
# Create a spatial version of census
census_spatial <- census
coordinates(census_spatial) <- ~x+y
proj4string(census_spatial) <- proj4string(moz)
# Switch to a projected format
# Project
census_spatial <- spTransform(census_spatial, CRS( "+init=epsg:3347" ) )
# Later, we can switch back to lat/long:
census_spatial_ll <- spTransform(census_spatial, CRS("+init=epsg:4326"))
```

# Clusters

In total, there are `r length(unique(village_df$cluster))` clusters.  Because there have been many changes in the "cleaning" process (ie, certain locations have apparently changed their village designations, etc.), the original clusters in some case overlap and are abnormally large. These clusters can be visualized below.

```{r, eval = TRUE}
possible_colors <- rainbow(1000)
plot(mop_ll)
title(main = 'All clusters')

clusters <- unique(sort(village_df$cluster))
for (i in 1:length(clusters)){
  this_cluster <- clusters[i]
  these_villages <- village_df$village_number[which(village_df$cluster == this_cluster)]
  these_houses <- census_spatial_ll[which(census_spatial_ll$village_number %in% these_villages),]
  # Get Convex hull
  border <- gConvexHull(these_houses)
  # Get a random color
  the_color <- sample(possible_colors, 1)
  plot(border, add = T,
       col = adjustcolor(the_color, alpha.f = 0.1),
       border = adjustcolor(the_color, alpha.f = 0.5))
  points(these_houses, pch = '.',
         col = adjustcolor(the_color, alpha.f = 0.8))
  message(this_cluster)
}
```

To help identify problematic clusters, each cluster's map is printed below.

```{r, eval = TRUE}
possible_colors <- rainbow(1000)
clusters <- unique(sort(village_df$cluster))
for (i in 1:length(clusters)){
  this_cluster <- clusters[i]
  plot(mop_ll)
  title(main = paste0('Cluster number: ', this_cluster))
  these_villages <- village_df$village_number[which(village_df$cluster == this_cluster)]
  title(sub = paste0('Includes villages :', paste0(these_villages, collapse = ', ')))
  these_houses <- census_spatial_ll[which(census_spatial_ll$village_number %in% these_villages),]
  # Get Convex hull
  border <- gConvexHull(these_houses)
  # Get a random color
  the_color <- sample(possible_colors, 1)
  plot(border, add = T,
       col = adjustcolor(the_color, alpha.f = 0.1),
       border = adjustcolor(the_color, alpha.f = 0.5))
  points(these_houses, pch = '.',
         col = adjustcolor(the_color, alpha.f = 0.8))
  message(this_cluster)
}
```

# Village table

The breakdown of villages can be viewed below.

```{r}
DT::datatable(village_df %>%
                dplyr::select(village_number,
                              cluster,
                              x_centroid,
                              y_centroid,
                              houses,
                              n_children))
```


# Cluster table

The breakdown of cluster assignments can be viewed below.

```{r}
x <- village_df %>%
  group_by(cluster) %>%
  summarise(village_numbers = paste0(village_number, collapse = ', '))

cluster_df <- census %>%
  group_by(cluster) %>%
  summarise(n_children = sum(number_of_children),
            houses_with_any_children = length(which(number_of_children > 0)),
            houses = n())

x <- x %>%
  left_join(cluster_df)
DT::datatable(x)

```

# Spray / no-spray table

The spray / no-spray designations were carried out in person. 


```{r}
# Read in the spray / no-spray statuses (emailed to me)
# spray_status <- read_excel('20160817 Mopeia spray-no spray random list.xlsx')
spray_status <- read_excel('Final spray no spray list.xlsx')

# Clean up
spray_status <- spray_status[,1:11]
names(spray_status) <-
  tolower(gsub('.', '_', names(spray_status), fixed = TRUE))
names(spray_status) <- gsub(' |-', '_', names(spray_status))

# Remove NA rows
spray_status <- 
  spray_status %>%
  filter(!is.na(cluster))

# Create a status column
spray_status$status <- spray_status$spray_status
spray_status$status <-
  ifelse(is.na(spray_status$status),
         FALSE,
         ifelse(spray_status$status == 'YES',
                TRUE,
                NA))

# Create a clusters dataframe
cluster_df <- census %>%
  group_by(cluster) %>%
  summarise(n_children = sum(number_of_children),
            houses_with_any_children = length(which(number_of_children > 0)),
            houses = n())


# Join to cluster_df
cluster_df$status <- NULL
cluster_df <- 
  cluster_df %>%
  left_join(spray_status %>%
              dplyr::select(cluster,
                            status))

```

The below table shows the spray / no-spray status _by cluster_.

```{r}
x <- cluster_df
DT::datatable(x %>% mutate(spray = ifelse(status, 'spray', 'no spray')) %>% dplyr::select(cluster,
                                                                                          spray))
```

The below table shows the spray / no-spray status _by village_.

```{r}
x <- 
  census %>%
  group_by(village_number) %>%
  summarise(houses = n(),
            cluster = first(cluster)) %>%
  left_join(cluster_df %>%
              dplyr::select(cluster, status))
DT::datatable(x %>% mutate(spray = ifelse(status, 'spray', 'no spray')) %>% dplyr::select(-status))
# Write a csv
# readr::write_csv(x, '~/Desktop/spray_no_spray_assignments.csv')
```


# Master table

The below table shows each location, with the corresponding village name, head of household name, health facility name, etc.

```{r}
# Bring the statuses into census
census <- census %>% 
  left_join(cluster_df %>%
              dplyr::select(cluster, status),
            by = 'cluster')

x <- census %>%
  mutate(spray = ifelse(status, 'spray', 'no spray')) %>%
  dplyr::select(cluster,
                village_number,
                spray,
                houseno_rev,
                head_name,
                lat,
                lng,
                health_facility) %>%
  arrange(cluster, village_number)
DT::datatable(x)
```

The spray / no-spray statuses can be visualized in the below maps. Blue is no spray; red is spray. Click to see details on each location.

# Map

## Spray / no-spray map

```{r}
# Create a spatial version of census
census_spatial <- census
coordinates(census_spatial) <- ~x+y
proj4string(census_spatial) <- proj4string(moz)
# Switch to a projected format
# Project
census_spatial <- spTransform(census_spatial, CRS( "+init=epsg:3347" ) )
# Later, we can switch back to lat/long:
census_spatial_ll <- spTransform(census_spatial, CRS("+init=epsg:4326"))

# # PLOT IT
# plot(mop_ll)
# cols <- ifelse(census$status == 'Spray', 'darkgreen', 'darkred')
# points(census_spatial_ll,
#        col = cols,
#        pch = '.')

# Divide into spray / no-spray
spray <- census %>% filter(status)
no_spray <- census %>% filter(!status)

leaflet() %>%
  addProviderTiles("Esri.WorldImagery") %>%
  addPolylines(data = mop_ll, color = 'black') %>%
  addCircleMarkers(lng = spray$lng,
  lat = spray$lat,
  color = colors,
  fillColor = 'red',
  radius = 2.5,
  opacity = 0,
  fillOpacity = 0.5,
  popup = paste0('Household: ', spray$houseno_rev, ' Village number: ',
  spray$village_number, ' Cluster number: ', spray$cluster)) %>%
  
  addCircleMarkers(lng = no_spray$lng,
  lat = no_spray$lat,
  color = colors,
  fillColor = 'blue',
  radius = 2.5,
  opacity = 0,
  fillOpacity = 0.5,
popup = paste0('Household: ', no_spray$houseno_rev, ' Village number: ',
  no_spray$village_number, ' Cluster number: ', no_spray$cluster)) 
```


## Cluster map

```{r}
the_colors <- colorRampPalette(brewer.pal(9, 'Spectral'))(length(unique(census$cluster)))
colors <- sample(the_colors, length(the_colors))
colors <- the_colors[as.numeric(factor(census$cluster))]
colors[is.na(colors)] <- 'white'
leaflet() %>%
  addProviderTiles("Esri.WorldImagery") %>%
  addPolylines(data = mop_ll, color = 'black') %>%
  addCircleMarkers(lng = census$lng,
  lat = census$lat,
  color = colors,
  fillColor = colors,
  radius = 2.5,
  opacity = 0,
  fillOpacity = 0.5,
  popup = paste0('Household: ', spray$houseno_rev, ' Village number: ',
  spray$village_number, ' Cluster number: ', spray$cluster))
```
