---
title: "Publication-quality visualizations for Mopeia COST study"
author: "Brew"
date: '`r Sys.Date()`'
output:
  html_document: default
  pdf_document: default
self_contained: no
subtitle: Reproducible code and output
link-citations: yes
---

What follows is the code used for the generation of publication quality figures. Note that certain dependencies' source code are not included in this document (for example, references to other `R` scripts); in these cases, all code is available at https://github.com/joebrew/zambezia.



```{r setup, include=FALSE}
library(tufte)
# invalidate cache when the tufte version changes
knitr::opts_chunk$set(tidy = FALSE, cache.extra = packageVersion('tufte'))
options(htmltools.dir.version = FALSE)

knitr::opts_chunk$set(comment = NA, 
               echo = TRUE, 
               warning = FALSE, 
               message = FALSE, 
               error = TRUE, 
               # fig.height = 5,
               # fig.width = 8,
               fig.align = 'center')
knitr::opts_chunk$set(
  fig.path = "images/"
)

```

```{r}
# Define theme for plotting
# Borrowed from https://rpubs.com/Koundy/71792
theme_publication <- function(base_size=14) {
  library(grid)
  library(ggthemes)
  (theme_foundation(base_size=base_size)
    + theme(plot.title = element_text(face = "bold",
                                      size = rel(1.2), hjust = 0.5),
            text = element_text(),
            panel.background = element_rect(colour = NA),
            plot.background = element_rect(colour = NA),
            panel.border = element_rect(colour = NA),
            axis.title = element_text(face = "bold",size = rel(1)),
            axis.title.y = element_text(angle=90,vjust =2),
            axis.title.x = element_text(vjust = -0.2),
            axis.text = element_text(), 
            axis.line = element_line(colour="black"),
            axis.ticks = element_line(),
            panel.grid.major = element_line(colour="#f0f0f0"),
            panel.grid.minor = element_blank(),
            legend.key = element_rect(colour = NA),
            legend.position = "bottom",
            legend.direction = "horizontal",
            legend.key.size= unit(0.5, "cm"),
            legend.margin = unit(0, "cm"),
            legend.title = element_text(face="italic"),
            plot.margin=unit(c(10,5,5,5),"mm"),
            strip.background=element_rect(colour="#f0f0f0",fill="#f0f0f0"),
            strip.text = element_text(face="bold")
    ))
}

colors <- colorRampPalette(c('orange', 'red', 'darkgrey'))(3)

scale_fill_publication <- function(...){
  library(scales)
  discrete_scale("fill",
                 "Publication",
                 manual_pal(values = colors
                              # c("#386cb0","#fdb462","#7fc97f","#ef3b2c","#662506","#a6cee3","#fb9a99","#984ea3","#ffff33")
                            ), ...)
  
}

scale_colour_publication <- function(...){
  library(scales)
  discrete_scale("colour",
                 "Publication",
                 manual_pal(values = colors
                            # c("#386cb0","#fdb462","#7fc97f","#ef3b2c","#662506","#a6cee3","#fb9a99","#984ea3","#ffff33")
                            ), ...)
  
}

compassRose<-function(x,y,rot=0,cex=1) {
  oldcex<-par(cex=cex)
  mheight<-strheight("M")
  xylim<-par("usr")
  plotdim<-par("pin")
  xmult<-(xylim[2]-xylim[1])/(xylim[4]-xylim[3])*plotdim[2]/plotdim[1]
  point.angles<-seq(0,7*pi/4,by=pi/4)+pi*rot/180
  crspans<-rep(c(mheight*3,mheight/2),4)
  xpoints<-cos(point.angles)*crspans*xmult+x
  ypoints<-sin(point.angles)*crspans+y
  polygon(xpoints,ypoints)
  txtxpoints<-cos(point.angles[c(1,3,5,7)])*1.33*crspans[1]*xmult+x
  txtypoints<-sin(point.angles[c(1,3,5,7)])*1.33*crspans[1]+y
  text(txtxpoints,txtypoints,c("E","N","W","S"))
  par(oldcex)
}

library(cism)
library(sp)
library(tidyverse)
library(broom)
library(maptools)
library(RColorBrewer)
library(ggrepel)
library(GISTools)  
library(maps)
```

# Figure 1: map of unidades sanitarias

The below is a map of all the health facilities in Mopeia. I have not included an "inset" map with the location of Mopeia in Mozambique because this was not possible (to fit without overlapping, and being so small as to be unusable, etc). Instead, I've included a separate map which can be manually pasted on as an inset, if so desired.

```{r, fig.width = 8, fig.height = 5}
mop2 <- cism::mop2
# us <- cism::us
# us <- us %>% filter(!is.na(longitude))
# coordinates(us) <- ~longitude+latitude
# proj4string(us) <- proj4string(mop2)
# us <- us[!is.na(over(us, polygons(mop2))),]

us <- read.csv('unidades_sanitarias_cleaned.csv')
# Make spatial
us$x <- us$lng <- us$lon <- us$longitude <- us$gpc_lng
us$y <- us$lat <- us$latitude <- us$gpc_lat
coordinates(us) <- ~gpc_lng+gpc_lat
proj4string(us) <- proj4string(mop2)
us <- us[!is.na(over(us, polygons(mop2))),]


plot(mop2)
maps::map.scale(x=35.3, y=-18.1, ratio=FALSE, relwidth=0.18)  
# north.arrow(xb=35.8, yb=-18.2, len=0.05, lab="N")  
compassRose(x=36.45, y=-18.15, cex = 0.5)
# points(us, pch = 1)
points(us, pch = "+")
legend('topleft',
       pch = "+",
       legend = 'Health\nfacilities',
       bty = 'n',
       cex = 0.9)

# par(usr=c(36.3, 36.4, -17., -17.7))
# rect(xleft =36.3,ybottom = -17.79 ,xright = 36.4,ytop = -17.73,col = "white")
# plot(moz3, add = TRUE)
# map(mop2,add=T)
```

# Figure 1b: possible inset map

```{r}
plot(moz0)
plot(mop2, col = 'darkred', add = TRUE, border = NA)
```


```{r bib, include=FALSE}
# create a bib file for the R packages used in this document
knitr::write_bib(c('base', 'rmarkdown'), file = 'skeleton.bib')
```

# Figure 2: Borders, etc.

```{r}
# Helpers
source('lib/get_zambezia.R')
source('lib/helpers.R')
```

```{r, echo = FALSE, warning = FALSE, message = FALSE, comment = NA, error= FALSE, cache = FALSE}
# No scientific notation
options(scipen=999)

# Packages 
library(xtable)
library(knitr)
library(ggplot2) 
# library(dplyr)
library(rgeos)
library(RColorBrewer)
library(Rmisc)
# Libraries
library(rgdal)
library(raster)
library(sp)
library(leaflet)
library(ggmap)
library(readxl)
library(dplyr)
library(readr)
library(geosphere)

# Basic knitr options
opts_chunk$set(comment = NA, 
               echo = FALSE, 
               warning = FALSE, 
               message = FALSE, 
               error = TRUE, # Render report, even with errors
               cache = F)

# Helpers
source('lib/get_zambezia.R')
source('lib/helpers.R')

#### READ IN AND CLEAN UP DATA

# Read in final dataset, as given by Orvahlo on October 24, 2016
# census <- read_excel('Limpesa_Final__241016_For maps.xlsx')
# Read in corrected final dataset, sent by Edgar on October 30, 2016
# census <- read_excel('Casos problematticos resolvidos_Edgar Jamisse_Mopeia_2016_28_10.xls')
census <- read_csv('Casos problematticos resolvidos_Edgar Jamisse_Mopeia_2016_28_10.csv')
# census <- read_csv('CensusMopeia_DATA_2016-11-11_1355.csv')
# Fix weird header
names(census)[6] <- 'locality_Final'

# Get more nicely named geographic coordinate column names
census$lat <- census$y <- census$latitude <- census$gpc_lat
census$lon <- census$lng <- census$longitude <-census$x <- census$gpc_lng

# Get a shapefile for Mozambique
moz <- raster::getData('GADM', country = 'MOZ', level = 3)
moz2 <- raster::getData('GADM', country = 'MOZ', level = 2)

# GET CORRECT FINAL COLUMNS, AS EXPLAINED IN EMAIL
# Get a village number
census$village_number <- census$`village number_final`
census$administrative_post <- census$administrative_post_final
census$house_number <- census$houseno_Final_1
census$locality <- census$locality_Final

# Convert village number to numeric
census$village_number <- 
  as.numeric(as.character(census$village_number))

# Get the corrected locality and bairro names sent by edgar on nov 7
locality_corrections <- read_excel('Mapa Actualizado de bairros Censados_07_11-2016_Edgar-Revisao_FIN.xlsx', skip = 2)

# Join to the census
x <- 
  census %>%
  left_join(locality_corrections,
            by = c('village_number' = 'CODIGO')) %>%
  mutate(BAIRRO = toupper(BAIRRO))
# Make updates
x <- x %>%
  mutate(administrative_post = 
           ifelse(administrative_post != `POSTO ADMIN` & !is.na(`POSTO ADMIN`), 
                  `POSTO ADMIN`, administrative_post),
         locality = ifelse(locality != LOCALIDADE & !is.na(LOCALIDADE), 
                           LOCALIDADE, locality),
         local_village_name = ifelse(local_village_name != BAIRRO & !is.na(BAIRRO),
                         BAIRRO, local_village_name)) %>%
  dplyr::select(-`POSTO ADMIN`, LOCALIDADE, BAIRRO)
# Overwite census
census <- x
rm(x)
# Remove all NA rows
census <- census %>%
  filter(!is.na(lng),
         !is.na(lat),
         !is.na(village_number),
         !is.na(house_number))


# Save old version of census
census_old <- census

# Create a spatial version of census
census_spatial <- census
coordinates(census_spatial) <- ~x+y
proj4string(census_spatial) <- proj4string(moz)

# Subset the shapefile just to Zambezia
zam <- moz[moz@data$NAME_1 == 'Zambezia',]

# Subset further to just Mopeia
mop <- moz2[moz2@data$NAME_2 == 'Mopeia',]

# Save the original proj4string
original_proj4string <- proj4string(moz)

# Prepare ggmap-compatible objects
zam_fortified <- fortify(zam, id = ID_3)
mop_fortified <- fortify(mop, id = ID_3)
zam_gg <- get_map(location = c(lon = mean(census$lon, na.rm = TRUE),
                               lat = mean(census$lat, na.rm = TRUE)),
                  maptype = 'satellite')
mop_gg <- get_map(location = c(lon = mean(census$lon, na.rm = TRUE),
                               lat = mean(census$lat, na.rm = TRUE)),
                  maptype = 'satellite')

# Switch to a projected format
# Project
census_spatial <- spTransform(census_spatial, CRS( "+init=epsg:3347" ) )
mop <- spTransform(mop, CRS("+init=epsg:3347"))
# Later, we can switch back to lat/long:
census_spatial_ll <- spTransform(census_spatial, CRS("+init=epsg:4326"))
mop_ll  <- spTransform(mop, CRS("+init=epsg:4326"))
# proj4string(mop_ll) <- proj4string(zam)

# Read in data at the child-specific level
children <- read_csv('redcap/CensusMopeia_DATA_2016-11-01_1734.csv')

# Keep only those households with children
children <- 
  children %>%
  filter(!is.na(perm_id_final)) %>%
  mutate(permid = perm_id_final) %>%
  dplyr::select(-perm_id_final)

# Keep only those children which appear in the non-problematic census records
children <-
  children %>%
  filter(parent_auri %in% census$uri) 

# Keep only the columns of interest
children <- 
  children %>%
  dplyr::select(permid,
                dob,
                name,
                gender,
                keeper,
                parent_auri)

# Get the household id
children <-
  children %>%
  left_join(census %>% 
              dplyr::select(house_number,
                            village_number,
                            uri),
            by = c('parent_auri' = 'uri'))

if('clean_maps_with_buffers_data.RData' %in% dir()){
  load('clean_maps_with_buffers_data.RData')
} else {
  # Get spray status (will use for clusters too)
  get_spray_status() # produces spray_village (village level) and spray_status (cluster level)
  
  # Make a dataframe of just villages (with cluster assignments)
  village_df <- make_village_df(census = census,
                                spray_village = spray_village)
  
  # Get a distance matrix (takes 2-3 minutes)
  distance_matrix <- get_distance_matrix(spatial_ll_census = census_spatial_ll)
  
  # Identify problems
  identify_problems()
  census$problematic <- census$house_number %in% incorrect_houses
  census_spatial$problematic <- census_spatial$house_number %in% incorrect_houses
  census_spatial_ll$problematic <- census_spatial_ll$house_number %in% incorrect_houses
  
  # SUBSET TO REMOVE PROBLEMATIC
  # Get distance matrix after subset
  distance_matrix_old <- distance_matrix
  distance_matrix <- distance_matrix[!census$problematic,
                                         !census$problematic]
  
  # Identify the nearest neighbors of each house
  nn <- nearest_neighbor(census_spatial_ll = census_spatial_ll,
                             distance_matrix_old = distance_matrix_old,
                             n = 5)
  census$nearest_neighbors <- nn
  census_old$nearest_neighbors <- nn
  census_spatial$nearest_neighbors <- nn
  census_spatial_ll$nearest_neighbors <- nn
  
  # Remove incorrect houses
  census <- census[which(!census$house_number %in% incorrect_houses),]
  census_spatial <- census_spatial[which(!census_spatial$house_number %in% incorrect_houses),]
  census_spatial_ll <- census_spatial_ll[which(!census_spatial_ll$house_number %in% incorrect_houses),]
  
  village_df <- village_df %>% filter(!is.na(village_number),
                                      !duplicated(village_number))

  # Join spray status to objects
  joiner <- 
    village_df %>%
                dplyr::select(village_number, cluster, status)
  
  census_spatial_ll@data <-
    left_join(census_spatial_ll@data,
              joiner,
              by = 'village_number')
  census_spatial@data <-
    left_join(census_spatial@data,
              joiner,
              by = 'village_number')
  census <-
    left_join(census,
              joiner,
              by = 'village_number')
  
  # Get the voronoi tesselation surface
  vv <- voronoi(spatial_census_ll = census_spatial_ll)
  # Get the voronoi tesselation surface at the level of the spray zones
  vv_spray <- voronoi_spray_level(spatial_census_ll = census_spatial_ll)
  
  # Get a non lat long version of voronoi surface
  # For the village level
  proj4string(vv) <- proj4string(census_spatial_ll)
  vv_ll <- vv
  vv <- sp::spTransform(vv, CRSobj = proj4string(census_spatial))
  
  # For the spray-level
  proj4string(vv_spray) <- proj4string(census_spatial_ll)
  vv_spray_ll <- vv_spray
  vv_spray <- sp::spTransform(vv_spray, CRSobj = proj4string(census_spatial))
    
  # Get the buffered voronoi tesselation surface
  vvb <- rgeos::gBuffer(vv,  # village level
                        byid = TRUE, 
                        id = vv$village_number,
                        width = -1000)
  vvb_spray <- rgeos::gBuffer(vv_spray,  # village level
                        byid = TRUE, 
                        id = vv$status,
                        width = -1000)
  
  # Convert back to lat lon
  vvb_ll <- spTransform(vvb, CRSobj = proj4string(census_spatial_ll))
  vvb_ll_spray <- spTransform(vvb_spray, CRSobj = proj4string(census_spatial_ll))
  
  # Identify buffers voronoi
  census_spatial_ll <- 
    identify_buffers_voronoi(
    census_spatial_ll = census_spatial_ll,
           distance_matrix = distance_matrix,
           vvb_ll = vvb_ll)
  
  # Identify buffers voronoi at spray level
  census_spatial_ll <- 
    identify_buffers_voronoi_spray_level(census_spatial_ll = census_spatial_ll,
           distance_matrix = distance_matrix,
           vvb_ll_spray = vvb_ll_spray)
  
  # Get buffers through distance to other villages
  census_spatial_ll <- identify_buffers(spatial_ll_census = census_spatial_ll,
                       distances = distance_matrix)
  
  # Update other objects with variables only in census_spatial_ll
  update_other_objects()

  # Save
  save.image('clean_maps_with_buffers_data.RData')  
}


# Set all NA statuses to FALSE
census$status[is.na(census$status)] <- FALSE
census_spatial$status[is.na(census_spatial$status)] <- FALSE
census_spatial_ll$status[is.na(census_spatial_ll$status)] <- FALSE
village_df$status[is.na(village_df$status)] <- FALSE

# Resource helpers
source('lib/helpers.R')

# Get children's spray status
children <- 
  left_join(children,
            census %>% 
              dplyr::select(uri,
                            status),
            by = c('parent_auri' = 'uri'))

# Remove vv_ll's out of district areas
x <- rgeos::gIntersection(vv_ll,
                        mop_ll,
                        byid = TRUE)
the_data <- vv@data
x <- SpatialPolygonsDataFrame(Sr = x,
                              data = the_data,
                              match.ID = FALSE)

# Overwrite
vv_ll <- x

# Remove vvb_ll's out of district areas
x <- rgeos::gIntersection(vvb_ll,
                        mop_ll,
                        byid = TRUE)
the_data <- vvb@data
x <- SpatialPolygonsDataFrame(Sr = x,
                              data = the_data,
                              match.ID = FALSE)

# Overwrite
vvb_ll <- x

# Remove vvb_ll_spray's out of district areas
x <- rgeos::gIntersection(vvb_ll_spray,
                        mop_ll,
                        byid = TRUE)
the_data <- vvb_ll_spray@data
x <- SpatialPolygonsDataFrame(Sr = x,
                              data = the_data,
                              match.ID = FALSE)

# Overwrite
vvb_ll_spray <- x


# Remove vvb_spray's out of district areas
x <- rgeos::gIntersection(vvb_spray,
                        mop,
                        byid = TRUE)
the_data <- vvb_spray@data
x <- SpatialPolygonsDataFrame(Sr = x,
                              data = the_data,
                              match.ID = FALSE)

# Overwrite
vvb_spray <- x


# Remove vv_spray_ll's out of district areas
x <- rgeos::gIntersection(vv_spray_ll,
                        mop_ll,
                        byid = TRUE)
the_data <- vv_spray_ll@data
x <- SpatialPolygonsDataFrame(Sr = x,
                              data = the_data,
                              match.ID = FALSE)

# Overwrite
vv_spray_ll <- x


# Remove those with no spray status
children <- 
  children %>%
  filter(!is.na(status))
cols <- colorRampPalette(brewer.pal(9, 'Spectral'))(nrow(vv_ll))
cols <- sample(cols, length(cols))
cols <- adjustcolor(cols, alpha.f = 0.9)

```

## Without colors (no buffer)

```{r}
plot(mop_ll)
plot(vv_ll, col = 'white', add = TRUE, border = adjustcolor('black', alpha.f = 0.5))
maps::map.scale(x=35.3, y=-18.1, ratio=FALSE, relwidth=0.18)  
# north.arrow(xb=35.8, yb=-18.2, len=0.05, lab="N")  
compassRose(x=36.45, y=-18.15, cex = 0.5)
```

## With colors (no buffer)

```{r}
plot(mop_ll)
plot(vv_ll, col = cols,add = TRUE, border = adjustcolor('black', alpha.f = 0.5))
plot(mop_ll, add = TRUE)
maps::map.scale(x=35.3, y=-18.1, ratio=FALSE, relwidth=0.18)  
# north.arrow(xb=35.8, yb=-18.2, len=0.05, lab="N")  
compassRose(x=36.45, y=-18.15, cex = 0.5)
```

We can apply our internal buffers and we get the following:

```{r}
cols <- colorRampPalette(brewer.pal(9, 'Spectral'))(nrow(vvb_ll))
cols <- sample(cols, length(cols))
cols <- adjustcolor(cols, alpha.f = 0.9)
```

## Without colors (buffer)

```{r}
plot(mop_ll)
plot(vvb_ll, col = 'white' ,add = TRUE, border = adjustcolor('black', alpha.f = 0.5))
plot(mop_ll, add = TRUE)
maps::map.scale(x=35.3, y=-18.1, ratio=FALSE, relwidth=0.18)  
# north.arrow(xb=35.8, yb=-18.2, len=0.05, lab="N")  
compassRose(x=36.45, y=-18.15, cex = 0.5)
```

## With colors (buffer)

```{r}
plot(mop_ll)
plot(vvb_ll, col = cols ,add = TRUE, border = adjustcolor('black', alpha.f = 0.5))
plot(mop_ll, add = TRUE)
maps::map.scale(x=35.3, y=-18.1, ratio=FALSE, relwidth=0.18)  
# north.arrow(xb=35.8, yb=-18.2, len=0.05, lab="N")  
compassRose(x=36.45, y=-18.15, cex = 0.5)
```

## With houses too (no colors)

```{r}
color_df <- data_frame(village_number = vvb_ll$village_number,
                       col = cols)
the_points <- census_spatial_ll 
the_points@data <- the_points@data %>%
  left_join(color_df,
            by = 'village_number')
```

```{r}
plot(mop_ll)
plot(vvb_ll, col = adjustcolor('white', alpha.f = 0.6),add = TRUE, border = 'black')
plot(mop_ll, add = TRUE)
points(the_points,
       col = adjustcolor('black', alpha.f = 0.2),
       pch = 3,
       cex = 0.2)
maps::map.scale(x=35.3, y=-18.1, ratio=FALSE, relwidth=0.18)  
# north.arrow(xb=35.8, yb=-18.2, len=0.05, lab="N")  
compassRose(x=36.45, y=-18.15, cex = 0.5)
legend(x = 35.47, y = -17.445,
       pch = 3,
       pt.cex = 0.5,
       col = c('black'),
       legend = c('House'),
       bty= 'n',
       cex = 0.85)
```

## With houses too (with colors)

```{r}
plot(mop_ll)
plot(vvb_ll, col = adjustcolor(cols, alpha.f = 0.6),add = TRUE, border = NA)
plot(mop_ll, add = TRUE)
points(the_points,
       col = adjustcolor(the_points@data$col, alpha.f = 0.5),
       pch = 3,
       cex = 0.5)
maps::map.scale(x=35.3, y=-18.1, ratio=FALSE, relwidth=0.18)  
# north.arrow(xb=35.8, yb=-18.2, len=0.05, lab="N")  
compassRose(x=36.45, y=-18.15, cex = 0.5)
legend(x = 35.47, y = -17.445,
       pch = 3,
       pt.cex = 0.5,
       col = c('black'),
       legend = c('House'),
       bty= 'n',
       cex = 0.85)
```

## Figure 3

Our spray / no-spray map looks like this:

```{r}
the_data <- vv_spray_ll
the_colors <- ifelse(the_data@data$status, 'darkred', 'darkgreen')
the_colors <- adjustcolor(the_colors, alpha.f = 0.6)
status <- the_data$status
plot(mop_ll)
plot(the_data, add = TRUE, col = the_colors, border = adjustcolor('black', alpha.f = 0.5))
legend(x = 36.5, y = -17.5,
       fill = the_colors,
       legend = c('Spray', 'No spray'),
       cex = 0.8,
       bty = 'n')
maps::map.scale(x=35.3, y=-18.1, ratio=FALSE, relwidth=0.18)  
# north.arrow(xb=35.8, yb=-18.2, len=0.05, lab="N")  
compassRose(x=36.45, y=-18.15, cex = 0.5)
```

We add 1 kilometer internal buffers to each border to get the following:

```{r}
the_data <- vvb_ll_spray
the_colors <- ifelse(the_data@data$status, 'darkred', 'darkgreen')
the_colors <- adjustcolor(the_colors, alpha.f = 0.6)
status <- the_data$status
plot(mop_ll)
plot(the_data, col = the_colors, border = adjustcolor('black', alpha.f = 0.5), add = TRUE)
legend(x = 36.5, y = -17.5,
       fill = the_colors,
       legend = c('Spray', 'No spray'),
       cex = 0.8,
       bty = 'n')
maps::map.scale(x=35.3, y=-18.1, ratio=FALSE, relwidth=0.18)  
# north.arrow(xb=35.8, yb=-18.2, len=0.05, lab="N")  
compassRose(x=36.45, y=-18.15, cex = 0.5)
```

We can then add each household onto our surface to visualize in which areas (spray, no-spray, or buffer) each households falls into.

```{r}
the_data <- vvb_ll_spray
the_colors <- ifelse(the_data@data$status, 'darkred', 'darkgreen')
the_colors <- adjustcolor(the_colors, alpha.f = 0.6)

plot(mop_ll)
plot(the_data, col = the_colors, border = adjustcolor('black', alpha.f = 0.5), add = TRUE)

legend(x = 36.5, y = -17.47,
       fill = the_colors,
       legend = c('Spray', 'No spray'),
       cex = 0.8,
       bty = 'n',
       title = 'Zones')
maps::map.scale(x=35.3, y=-18.1, ratio=FALSE, relwidth=0.18)  
# north.arrow(xb=35.8, yb=-18.2, len=0.05, lab="N")  
compassRose(x=36.45, y=-18.15, cex = 0.5)

points(census_spatial_ll,
       col = ifelse(census_spatial_ll$within_1k_voronoi_buffer_spray, 
                    'blue',
                    ifelse(census_spatial_ll$status,
                           'darkred',
                           ifelse(!census_spatial_ll$status,
                                  'darkgreen',
                                  ifelse(is.na(census_spatial_ll$status), 'darkgreen', NA)))),
       pch = 3,
       cex = 0.3)
legend(x = 35.47, y = -17.445,
       pch = 3,
       pt.cex = 0.5,
       col = c('darkred', 'darkgreen', 'blue'),
       legend = c('Spray', 'No spray', 'Buffer'),
       bty= 'n',
       cex = 0.65,
       title = 'Houses')
```

